    \expectedproblemsection{Standard Mathematical Problems}

    \begin{expectedproblem}[1988 IMO]
        Let $a$ and $b$ be positive integers such that $ab+1$ divides
        $a^2+b^2$. Show that
        \[
            \frac{a^2+b^2}{ab+1}
        \]
        is the square of an integer.
    \end{expectedproblem}

    \begin{solution}
        Let
        \[
            k=\frac{a^2+b^2}{ab+1}\in\N.
        \]
        Suppose, for contradiction, that $k$ is not a perfect square. Among all
        positive integer solutions of
        \[
            a^2+b^2=k(ab+1),
        \]
        choose one for which $a+b$ is minimal, and assume without loss of
        generality that $a\geq b$.

        If $a=b$, then
        \[
            k=\frac{2a^2}{a^2+1}<2,
        \]
        so $k=1$, a contradiction. Hence $a>b$. Regard the equation as a
        quadratic equation in $a$:
        \[
            x^2-kbx+b^2-k=0.
        \]
        One root is $a$, so the other root is
        \[
            c=kb-a\in\Z.
        \]
        We claim that $c>0$. From
        \[
            a(a-kb)=k-b^2,
        \]
        if $a=kb$, then $k=b^2$, contrary to the assumption. If $a>kb$, then
        the left-hand side is at least $a\geq kb\geq k$, whereas the
        right-hand side is less than $k$, which is impossible. Therefore
        $a<kb$, and hence $c>0$.

        Since $k$ is a positive integer that is not a perfect square, we have
        $k\geq2$. Let
        \[
            f(x)=x^2-kbx+b^2-k.
        \]
        Since
        \[
            f(b)=(2-k)b^2-k<0
        \]
        and the two roots are $c$ and $a$ with $a>b$, it follows that
        \[
            0<c<b.
        \]
        By Vi\`ete's formulas, $c$ is the other root, and therefore
        \[
            c^2+b^2=k(cb+1).
        \]
        Thus $(c,b)$ is another positive integer solution with
        \[
            c+b<a+b,
        \]
        contradicting the minimality of $a+b$. Hence $k$ must be a perfect
        square.
    \end{solution}

    \expectedproblemsection{Direct Calculation Problems}

    \expectedproblemsection{Memorization Problems}

    \expectedproblemsection{Reasoning and Puzzle Problems}

    \begin{expectedproblem}[Happy Ending Problem]
        Five points are given in the plane, no three of which are collinear.
        Prove that four of them are the vertices of a convex quadrilateral.
    \end{expectedproblem}

    \begin{solution}
        Consider the convex hull of the five points. If the convex hull has at
        least four vertices, any four of its vertices, taken in cyclic order,
        form a convex quadrilateral.

        It remains to consider the case in which the convex hull is a triangle
        with vertices $A,B,C$. Let the remaining two points be $P$ and $Q$;
        both lie in the interior of $\triangle ABC$. The line $PQ$ contains none
        of $A,B,C$, so two of the three vertices lie in the same open half-plane
        determined by $PQ$. Relabel them as $A$ and $B$.

        The triangle $\triangle ABQ$ meets the line $PQ$ only at $Q$, because
        $A$ and $B$ lie strictly on the same side of that line. Hence
        $P\notin\triangle ABQ$. Similarly, $Q\notin\triangle ABP$. Since
        $A$ and $B$ are vertices of the convex hull of all five points, neither
        lies in the triangle determined by the other three points. Consequently,
        all four points $A,B,P,Q$ are vertices of their convex hull, which is a
        convex quadrilateral.
    \end{solution}

    \begin{expectedproblem}[Freudenthal's Sum-and-Product Problem, Impossible Puzzle]
        Two integers $x$ and $y$ satisfy
        \[
            1<x<y,
            \qquad
            x+y\leq100.
        \]
        The sum $x+y$ is disclosed privately to $S$, and the product $xy$ is
        disclosed privately to $P$. The preceding information is common
        knowledge. The following conversation then takes place:
        \begin{enumerate}
            \item[(i)] $P$: ``I do not know the pair.''
            \item[(ii)] $S$: ``I knew that you did not know the pair.''
            \item[(iii)] $P$: ``Now I know the pair.''
            \item[(iv)] $S$: ``Now I also know the pair.''
        \end{enumerate}
        Determine $(x,y)$.
    \end{expectedproblem}

    \begin{solution}
        Let
        \[
            \Omega_0
            =
            \{(x,y)\in\N^2\mid (1<x<y)\land(x+y\leq100)\}.
        \]
        For $\Omega\subseteq\Omega_0$ and $\omega=(x,y)\in\Omega$, define the
        information sets
        \[
            \mathcal{I}_S^\Omega(\omega)
            =
            \{(u,v)\in\Omega\mid u+v=x+y\},
        \]
        \[
            \mathcal{I}_P^\Omega(\omega)
            =
            \{(u,v)\in\Omega\mid uv=xy\}.
        \]
        The four announcements successively restrict the state space as follows:
        \[
            \Omega_1
            =
            \{\omega\in\Omega_0\mid
              |\mathcal{I}_P^{\Omega_0}(\omega)|>1\},
        \]
        \[
            \Omega_2
            =
            \{\omega\in\Omega_1\mid
              \mathcal{I}_S^{\Omega_0}(\omega)\subseteq\Omega_1\},
        \]
        \[
            \Omega_3
            =
            \{\omega\in\Omega_2\mid
              |\mathcal{I}_P^{\Omega_2}(\omega)|=1\},
        \]
        and
        \[
            \Omega_4
            =
            \{\omega\in\Omega_3\mid
              |\mathcal{I}_S^{\Omega_3}(\omega)|=1\}.
        \]

        The finite elimination is summarized by
        \[
            \begin{array}{c|l|c}
                \text{stage} & \text{public information retained} &
                \text{number of states}\\
                \hline
                0 & (1<x<y)\land(x+y\leq100) & 2352\\
                1 & P\text{ does not know the pair} & 1747\\
                2 & S\text{ knew that }P\text{ did not know} & 145\\
                3 & P\text{ now knows the pair} & 86\\
                4 & S\text{ now knows the pair} & 1
            \end{array}
        \]
        At the second stage, the possible sums are exactly
        \[
            11,\ 17,\ 23,\ 27,\ 29,\ 35,\ 37,\ 41,\ 47,\ 53.
        \]
        These values follow by checking, for each admissible sum, whether every
        compatible factorization has at least two representations in
        $\Omega_0$. After the third announcement, the only sum-information set
        of cardinality $1$ is
        \[
            \mathcal{I}_S^{\Omega_3}(4,13)=\{(4,13)\}.
        \]
        Therefore,
        \[
            \Omega_4=\{(4,13)\},
        \]
        and the unique pair is
        \[
            \boxed{(x,y)=(4,13)}.
        \]
    \end{solution}
