    \chapter{Mathematical Prizes and Congresses}

    Science Quiz often asks for the short association between a prize, a
    congress, a laureate, and a mathematical achievement. This chapter records
    the facts that should be recognized quickly: not full biographies, but the
    names, years, keywords, and historical landmarks most likely to become quiz
    answers.

    \begin{fact}[Science Quiz Prize Triad]
        The three mathematics prizes that should be recognized first are
        \[
            \text{Fields Medal},\qquad
            \text{Abel Prize},\qquad
            \text{Wolf Prize in Mathematics}.
        \]
        In a fast quiz setting, memorize each prize by the pattern
        \[
            \text{prize}
            \quad\longleftrightarrow\quad
            \text{laureate}
            \quad\longleftrightarrow\quad
            \text{achievement or area}.
        \]
    \end{fact}

    \begin{strategy}
        History questions often ask only for a short association, not a full
        biography. When a mathematician is currently prominent, recently awarded,
        or connected to a \POSTECH--\KAIST{}
        context, add that person to the season-specific checklist in
        Mathematics in 2026.
    \end{strategy}

    \section*{Fields Medal and ICM}

    The International Congress of Mathematicians, commonly abbreviated as ICM,
    is the most important international congress in mathematics. The Fields
    Medal is awarded by the International Mathematical Union at the ICM every
    four years. It is named after John Charles Fields. A Fields Medalist must
    satisfy the age condition: the candidate's $40$th birthday must not occur
    before January $1$ of the year of the Congress.

    \vspace{1em}

    {\small
    \renewcommand{\arraystretch}{1.18}
    \begin{longtable}{@{}L{0.07\textwidth}L{0.16\textwidth}L{0.33\textwidth}L{0.36\textwidth}@{}}
        \toprule
        Year & ICM location & Medalists & Quiz keywords \\
        \midrule
        \endfirsthead

        \toprule
        Year & ICM location & Medalists & Quiz keywords \\
        \midrule
        \endhead

        \midrule
        \multicolumn{4}{r}{Continued on the next page.} \\
        \endfoot

        \bottomrule
        \endlastfoot

        1936
        & Oslo, Norway
        & Lars Ahlfors; Jesse Douglas
        & Complex analysis, Riemann surfaces; minimal surfaces, Plateau problem. \\

        1950
        & Cambridge, USA
        & Laurent Schwartz; Atle Selberg
        & Distribution theory; analytic number theory, sieve methods, zeta functions. \\

        1954
        & Amsterdam, Netherlands
        & Kunihiko Kodaira; Jean-Pierre Serre
        & Complex geometry, algebraic geometry; algebraic topology, sheaf theory. \\

        1958
        & Edinburgh, UK
        & Klaus Roth; René Thom
        & Diophantine approximation; cobordism, differential topology. \\

        1962
        & Stockholm, Sweden
        & Lars Hörmander; John Milnor
        & Partial differential equations; differential topology, exotic spheres. \\

        1966
        & Moscow, USSR
        & Michael Atiyah; Paul Cohen; Alexander Grothendieck; Stephen Smale
        & K-theory and index theorem; forcing and continuum hypothesis;
          modern algebraic geometry; generalized Poincaré conjecture. \\

        1970
        & Nice, France
        & Alan Baker; Heisuke Hironaka; Serge Novikov; John G. Thompson
        & Transcendental number theory; resolution of singularities;
          topology and geometry; finite group theory. \\

        1974
        & Vancouver, Canada
        & Enrico Bombieri; David Mumford
        & Number theory, complex analysis, minimal surfaces; moduli spaces and
          algebraic geometry. \\

        1978
        & Helsinki, Finland
        & Pierre Deligne; Charles Fefferman; Grigory Margulis; Daniel Quillen
        & Weil conjectures; several complex variables; Lie groups and ergodic
          theory; algebraic K-theory. \\

        1982
        & Warsaw, Poland (held in 1983)
        & Alain Connes; William Thurston; Shing-Tung Yau
        & Noncommutative geometry; $3$-manifolds and low-dimensional topology;
          Calabi conjecture and geometric analysis. \\

        1986
        & Berkeley, USA
        & Simon Donaldson; Gerd Faltings; Michael Freedman
        & Gauge theory and $4$-manifolds; Mordell conjecture and arithmetic
          geometry; topological $4$-dimensional Poincaré conjecture. \\

        1990
        & Kyoto, Japan
        & Vladimir Drinfeld; Vaughan Jones; Shigefumi Mori; Edward Witten
        & Quantum groups and Langlands program; Jones polynomial; minimal
          model program; mathematical physics and quantum field theory. \\

        1994
        & Zurich, Switzerland
        & Jean Bourgain; Pierre-Louis Lions; Jean-Christophe Yoccoz; Efim Zelmanov
        & Analysis and Banach spaces; partial differential equations; dynamical
          systems; restricted Burnside problem. \\

        1998
        & Berlin, Germany
        & Richard Borcherds; W. Timothy Gowers; Maxim Kontsevich; Curtis T. McMullen
        & Moonshine and automorphic forms; functional analysis and combinatorics;
          algebraic geometry and mathematical physics; complex dynamics. \\

        2002
        & Beijing, China
        & Laurent Lafforgue; Vladimir Voevodsky
        & Langlands correspondence over function fields; motivic cohomology and
          $\mathbb{A}^1$-homotopy theory. \\

        2006
        & Madrid, Spain
        & Andrei Okounkov; Grigori Perelman; Terence Tao; Wendelin Werner
        & Probability, representation theory, algebraic geometry; Ricci flow;
          PDE, harmonic analysis, additive combinatorics; SLE and Brownian motion. \\

        2010
        & Hyderabad, India
        & Elon Lindenstrauss; Ngo Bao Chau; Stanislav Smirnov; Cédric Villani
        & Ergodic theory and number theory; Fundamental Lemma; conformal
          invariance; Boltzmann equation and optimal transport. \\

        2014
        & Seoul, South Korea
        & Artur Avila; Manjul Bhargava; Martin Hairer; Maryam Mirzakhani
        & Dynamical systems; geometry of numbers; stochastic PDE;
          Riemann surfaces and moduli spaces. \\

        2018
        & Rio de Janeiro, Brazil
        & Caucher Birkar; Alessio Figalli; Peter Scholze; Akshay Venkatesh
        & Fano varieties and minimal model program; optimal transport;
          $p$-adic geometry; number theory and homogeneous dynamics. \\

        2022
        & Virtual; awards in Helsinki, Finland
        & Hugo Duminil-Copin; June Huh; James Maynard; Maryna Viazovska
        & Phase transitions in statistical physics; Hodge theory in combinatorics
          and matroids; prime numbers; sphere packing and the $E_8$ lattice. \\
    \end{longtable}
    }

    \begin{fact}[Essential ICM and Fields Medal landmarks]
        The following associations are especially useful.
        \begin{itemize}
            \item ICM stands for International Congress of Mathematicians.
            \item IMU stands for International Mathematical Union.
            \item The Fields Medal was first awarded in 1936 at ICM Oslo.
            \item The first two Fields Medalists were Lars Ahlfors and Jesse Douglas.
            \item The first year with four Fields Medalists was 1966.
            \item Jean-Pierre Serre is the youngest Fields Medalist.
            \item Edward Witten was the first physicist to receive the Fields Medal; as an
            undergraduate, he majored in history rather than mathematics.
            \item Andrew Wiles received a special IMU silver plaque in 1998 for
                  his proof of Fermat's Last Theorem.
            \item Grigori Perelman declined the Fields Medal in 2006.
            \item Maryam Mirzakhani was the first woman Fields Medalist.
            \item Maryna Viazovska was the second woman Fields Medalist.
            \item ICM 2014 was held in Seoul, South Korea.
            \item ICM 2022 was held as a virtual congress; the prize ceremony was
                  held in Helsinki, Finland.
            \item ICM 2026 is scheduled to be held in Philadelphia, USA.
        \end{itemize}
    \end{fact}

    \section*{Abel Prize}

    The Abel Prize is awarded by the Norwegian Academy of Science and Letters.
    It is named after Niels Henrik Abel and was first awarded in 2003. Unlike
    the Fields Medal, the Abel Prize has no age restriction. For Science Quiz,
    the Abel Prize is often asked through the pattern
    \[
        \text{year} \quad \longleftrightarrow \quad \text{laureate} \quad
        \longleftrightarrow \quad \text{famous theorem or area}.
    \]

    \vspace{1em}

    {\small
    \renewcommand{\arraystretch}{1.18}
    \begin{longtable}{@{}L{0.08\textwidth}L{0.32\textwidth}L{0.52\textwidth}@{}}
        \toprule
        Year & Laureate(s) & Quiz keywords \\
        \midrule
        \endfirsthead

        \toprule
        Year & Laureate(s) & Quiz keywords \\
        \midrule
        \endhead

        \midrule
        \multicolumn{3}{r}{Continued on the next page.} \\
        \endfoot

        \bottomrule
        \endlastfoot

        2003 & Jean-Pierre Serre
        & Algebraic topology, algebraic geometry, number theory; first Abel Prize laureate. \\

        2004 & Michael Atiyah; Isadore Singer
        & Atiyah--Singer index theorem. \\

        2005 & Peter D. Lax
        & Partial differential equations and applied mathematics. \\

        2006 & Lennart Carleson
        & Harmonic analysis; convergence of Fourier series. \\

        2007 & Srinivasa S. R. Varadhan
        & Probability theory; large deviations. \\

        2008 & John G. Thompson; Jacques Tits
        & Finite groups, classification of finite simple groups; buildings and group theory. \\

        2009 & Mikhail Gromov
        & Geometry, geometric group theory, symplectic geometry. \\

        2010 & John Tate
        & Number theory and arithmetic geometry. \\

        2011 & John Milnor
        & Topology, geometry, dynamics; exotic spheres. \\

        2012 & Endre Szemerédi
        & Discrete mathematics and theoretical computer science; Szemerédi's theorem. \\

        2013 & Pierre Deligne
        & Algebraic geometry; Weil conjectures. \\

        2014 & Yakov Sinai
        & Dynamical systems, ergodic theory, mathematical physics. \\

        2015 & John F. Nash Jr.; Louis Nirenberg
        & Nonlinear PDE, geometry, Nash embedding theorem. \\

        2016 & Andrew Wiles
        & Fermat's Last Theorem; modularity of elliptic curves. \\

        2017 & Yves Meyer
        & Wavelet theory; applications from signal compression to gravitational-wave analysis. \\

        2018 & Robert P. Langlands
        & Langlands program; representation theory and number theory. \\

        2019 & Karen Uhlenbeck
        & Geometric analysis and gauge theory; first woman Abel Prize laureate. \\

        2020 & Hillel Furstenberg; Gregory Margulis
        & Probability, dynamics, ergodic theory, Lie groups. \\

        2021 & László Lovász; Avi Wigderson
        & Discrete mathematics, theoretical computer science, complexity theory. \\

        2022 & Dennis P. Sullivan
        & Topology, geometric topology, dynamical systems. \\

        2023 & Luis A. Caffarelli
        & Nonlinear partial differential equations, free-boundary problems, Monge--Ampère equation. \\

        2024 & Michel Talagrand
        & Probability theory, functional analysis, spin glasses, concentration inequalities. \\

        2025 & Masaki Kashiwara
        & Algebraic analysis, $D$-modules, representation theory, crystal bases. \\

        2026 & Gerd Faltings
        & Arithmetic geometry; Mordell conjecture, Lang's conjectures, Diophantine geometry. \\
    \end{longtable}
    }

    \begin{fact}[Essential Abel Prize landmarks]
        The following associations are especially likely to be useful.
        \begin{itemize}
            \item Abel Prize $\leftrightarrow$ Norwegian Academy of Science and Letters.
            \item Abel Prize $\leftrightarrow$ Niels Henrik Abel.
            \item First Abel Prize laureate $\leftrightarrow$ Jean-Pierre Serre.
            \item First woman Abel Prize laureate $\leftrightarrow$ Karen Uhlenbeck.
            \item In common multiple-choice quiz settings, the Abel Prize money is closest to about
            one million US dollars.
            \item John Milnor $\leftrightarrow$ Fields Medal 1962 and Abel Prize 2011.
            \item Andrew Wiles $\leftrightarrow$ Abel Prize 2016 $\leftrightarrow$ Fermat's Last
            Theorem.
            \item Yves Meyer $\leftrightarrow$ Abel Prize 2017 $\leftrightarrow$ wavelets.
            \item Robert Langlands $\leftrightarrow$ Abel Prize 2018 $\leftrightarrow$ Langlands
            program.
            \item Gerd Faltings $\leftrightarrow$ Fields Medal 1986 and Abel Prize 2026.
        \end{itemize}
    \end{fact}

    \section*{Wolf Prize in Mathematics}

    The Wolf Prize is awarded by the Wolf Foundation. In the sciences, the Wolf
    Prize is awarded in Medicine, Agriculture, Mathematics, Chemistry, and
    Physics. It is especially important historically because many Wolf Prize in
    Mathematics laureates later received, or had already received, other major
    mathematical prizes. Unlike the Fields Medal, it is not restricted by age.

    \vspace{1em}

    {\small
    \renewcommand{\arraystretch}{1.16}
    \begin{longtable}{@{}L{0.09\textwidth}L{0.38\textwidth}L{0.45\textwidth}@{}}
        \toprule
        Year & Laureate(s) & Quiz keywords \\
        \midrule
        \endfirsthead

        \toprule
        Year & Laureate(s) & Quiz keywords \\
        \midrule
        \endhead

        \midrule
        \multicolumn{3}{r}{Continued on the next page.} \\
        \endfoot

        \bottomrule
        \endlastfoot

        1978 & Israel Gelfand; Carl L. Siegel
        & Representation theory, functional analysis; number theory, several complex variables. \\

        1979 & Jean Leray; André Weil
        & PDE and algebraic topology; algebraic geometry and number theory. \\

        1980 & Henri Cartan; Andrey Kolmogorov
        & Several complex variables and algebraic topology; probability theory and dynamical systems. \\

        1981 & Lars Ahlfors; Oscar Zariski
        & Complex analysis and Riemann surfaces; algebraic geometry. \\

        1982 & Mark Krein; Hassler Whitney
        & Functional analysis; differential topology and Whitney embedding. \\

        1983 & Shiing-Shen Chern
        & Global differential geometry, Chern classes. \\

        1984 & Paul Erdős
        & Combinatorics, number theory, probabilistic method. \\

        1985 & Kunihiko Kodaira; Hans Lewy
        & Complex geometry; partial differential equations. \\

        1986 & Samuel Eilenberg; Atle Selberg
        & Algebraic topology and category theory; analytic number theory and Selberg trace formula. \\

        1987 & Kiyoshi Itô; Peter D. Lax
        & Stochastic calculus; partial differential equations and applied mathematics. \\

        1988 & Lars Hörmander; Friedrich Hirzebruch
        & Partial differential equations; topology, characteristic classes, Hirzebruch--Riemann--Roch. \\

        1989 & Alberto Calderón; John Milnor
        & Harmonic analysis and singular integrals; topology and exotic spheres. \\

        1990 & Ennio De Giorgi; Ilya Piatetski-Shapiro
        & Minimal surfaces and regularity theory; automorphic forms. \\

        1992 & Lennart Carleson; John G. Thompson
        & Harmonic analysis; finite group theory. \\

        1993 & Mikhail Gromov; Jacques Tits
        & Geometry and geometric group theory; buildings and group theory. \\

        1995/96 & Robert Langlands; Andrew Wiles
        & Langlands program; Fermat's Last Theorem. \\

        1996/97 & Joseph Keller; Yakov Sinai
        & Applied mathematics and waves; dynamical systems and mathematical physics. \\

        1999 & László Lovász; Elias M. Stein
        & Discrete mathematics and algorithms; harmonic analysis. \\

        2000 & Raoul Bott; Jean-Pierre Serre
        & Topology and geometry; algebraic topology, algebraic geometry, number theory. \\

        2001 & Vladimir Arnold; Saharon Shelah
        & Dynamical systems and singularity theory; mathematical logic and set theory. \\

        2002/03 & Mikio Sato; John Tate
        & Algebraic analysis and hyperfunctions; arithmetic geometry. \\

        2005 & Gregory Margulis; Serge Novikov
        & Lie groups, ergodic theory, discrete subgroups; topology and geometry. \\

        2006/07 & Stephen Smale; Hillel Furstenberg
        & Differential topology and dynamical systems; ergodic theory and probability. \\

        2008/09 & Pierre Deligne; Phillip Griffiths; David Mumford
        & Algebraic geometry, Hodge theory, moduli spaces. \\

        2010 & Shing-Tung Yau; Dennis Sullivan
        & Geometric analysis; topology and dynamics. \\

        2012 & Michael Aschbacher; Luis Caffarelli
        & Finite groups; nonlinear PDE and free-boundary problems. \\

        2013 & George Mostow; Michael Artin
        & Rigidity, geometry, Lie groups; algebraic geometry. \\

        2014 & Peter Sarnak
        & Number theory, analysis, geometry, automorphic forms. \\

        2015 & James Arthur
        & Trace formula, automorphic representations. \\

        2017 & Richard Schoen; Charles Fefferman
        & Geometric analysis; analysis and partial differential equations. \\

        2018 & Alexander Beilinson; Vladimir Drinfeld
        & Geometric representation theory, mathematical physics, Langlands program. \\

        2019 & Jean-François Le Gall; Gregory Lawler
        & Probability theory, random walks, Brownian motion, SLE. \\

        2020 & Simon Donaldson; Yakov Eliashberg
        & Differential geometry, topology, gauge theory; symplectic and contact topology. \\

        2022 & George Lusztig
        & Representation theory and related areas. \\

        2023 & Ingrid Daubechies
        & Wavelet theory, time-frequency analysis, signal processing. \\

        2024 & Adi Shamir; Noga Alon
        & Mathematical cryptography; combinatorics, graph theory, theoretical computer science. \\
    \end{longtable}
    }

    \begin{fact}[Essential Wolf Prize landmarks]
        The following associations are the most useful ones to memorize first.
        \begin{itemize}
            \item Wolf Prize $\leftrightarrow$ Wolf Foundation.
            \item Wolf Prize in Mathematics $\leftrightarrow$ no Fields-style age restriction.
            \item Paul Erdős $\leftrightarrow$ Wolf Prize 1984 $\leftrightarrow$ combinatorics and
            probabilistic method.
            \item John Milnor $\leftrightarrow$ Wolf Prize 1989; together with the Fields Medal and
            Abel Prize, this makes him a standard example connected to all three major mathematics
            prizes.
            \item Andrew Wiles and Robert Langlands shared the 1995/96 Wolf Prize in Mathematics.
            \item Ingrid Daubechies $\leftrightarrow$ Wolf Prize 2023 $\leftrightarrow$ wavelets.
            \item Adi Shamir and Noga Alon $\leftrightarrow$ Wolf Prize 2024 $\leftrightarrow$
            cryptography and combinatorics.
        \end{itemize}
    \end{fact}

    \begin{fact}[Fast prize recognition]
        For quick Science Quiz recall, remember the following map first.
        \begin{itemize}
            \item Fields Medal $\leftrightarrow$ ICM $\leftrightarrow$ IMU $\leftrightarrow$ every
            four years $\leftrightarrow$ age condition.
            \item Abel Prize $\leftrightarrow$ Norwegian Academy of Science and Letters
            $\leftrightarrow$ no age restriction.
            \item Wolf Prize $\leftrightarrow$ Wolf Foundation $\leftrightarrow$ no Fields-style age
            restriction.
            \item Fields Medal questions often ask for the year, ICM location, laureate, and
            keyword.
            \item Abel and Wolf questions often ask for the laureate--area association.
        \end{itemize}
    \end{fact}

    \chapter{Famous Problems and Conjectures}

    Famous problems are usually not tested by asking for complete proofs. They
    are tested through the following information: the standard statement, the
    person or group attached to the problem, whether it is solved, and the main
    idea or keyword attached to the solution. The goal of this chapter is to
    make those associations immediate.

    \section*{Millennium Prize Problems}

    The Clay Mathematics Institute announced the seven Millennium Prize Problems
    in 2000. Each problem carries a prize of one million US dollars. Among the
    seven, only the Poincaré conjecture has been solved.

    \vspace{1em}

    {\small
    \renewcommand{\arraystretch}{1.18}
    \begin{longtable}{@{}L{0.24\textwidth}L{0.30\textwidth}L{0.16\textwidth}L{0.22\textwidth}@{}}
        \toprule
        Problem & Standard mathematical form & Status & Quiz anchors \\
        \midrule
        \endfirsthead

        \toprule
        Problem & Standard mathematical form & Status & Quiz anchors \\
        \midrule
        \endhead

        \midrule
        \multicolumn{4}{r}{Continued on the next page.} \\
        \endfoot

        \bottomrule
        \endlastfoot

        Poincaré Conjecture
        & Every closed simply connected $3$-manifold is homeomorphic to $S^3$.
        & Solved.
        & Poincaré; Perelman; Ricci flow. \\

        Riemann Hypothesis
        & Every nontrivial zero of $\zeta(s)$ has real part $1/2$.
        & Open.
        & Riemann; zeta function; prime numbers. \\

        P versus NP
        & Is $P=NP$?
        & Open.
        & Cook, Karp; computational complexity. \\

        Birch and Swinnerton-Dyer Conjecture
        & For an elliptic curve $E$, $\rk E(\Q)$ equals the order of
          vanishing of $L(E,s)$ at $s=1$.
        & Open.
        & Elliptic curves; $L$-functions. \\

        Hodge Conjecture
        & Rational Hodge classes on smooth projective complex varieties should
          come from algebraic cycles.
        & Open.
        & Hodge theory; algebraic geometry. \\

        Navier--Stokes Existence and Smoothness
        & Smooth initial data for the $3$-dimensional incompressible
          Navier--Stokes equations should have global smooth solutions, or one
          must find a breakdown.
        & Open.
        & Fluid mechanics; PDE; turbulence. \\

        Yang--Mills Existence and Mass Gap
        & Construct quantum Yang--Mills theory on $\R^4$ for compact simple gauge
          groups and prove a positive mass gap.
        & Open.
        & Gauge theory; quantum field theory; mass gap. \\
    \end{longtable}
    }

    \begin{fact}[Millennium Prize status check]
        Among the seven Millennium Prize Problems, the Poincaré conjecture is
        the only one that has been solved. The remaining six are the Riemann
        Hypothesis, P versus NP, the Birch and Swinnerton-Dyer conjecture, the
        Hodge conjecture, Navier--Stokes existence and smoothness, and
        Yang--Mills existence and mass gap.
    \end{fact}

    \begin{theorem}[Poincaré Conjecture]
        Every closed simply connected $3$-manifold is homeomorphic to the
        $3$-sphere $S^3$.
    \end{theorem}

    \begin{remark}
        Henri Poincaré posed the problem in the early twentieth century. Grigori
        Perelman proved it using Ricci flow and ideas from Hamilton's program.
        The key Science Quiz association is
        \[
            \text{Poincaré conjecture}
            \quad\longleftrightarrow\quad
            \text{Perelman}
            \quad\longleftrightarrow\quad
            \text{Ricci flow}.
        \]
    \end{remark}

    \begin{hypothesis}[Riemann Hypothesis]
        Let
        \[
            \zeta(s)=\sum_{n=1}^{\infty}\frac{1}{n^s}
        \]
        be the Riemann zeta function, initially defined for $\Real(s)>1$ and
        continued meromorphically. The Riemann Hypothesis states that every
        nontrivial zero of $\zeta(s)$ has real part $1/2$.
    \end{hypothesis}

    \begin{idea}
        The zeta function is connected to prime numbers by Euler's product
        \[
            \zeta(s)=\prod_{p}\frac{1}{1-p^{-s}}.
        \]
        Thus RH is a statement about the hidden regularity of the distribution
        of prime numbers.
    \end{idea}

    \begin{openproblem}[P versus NP]
        Determine whether
        \[
            P=NP.
        \]
        Here $P$ is the class of decision problems solvable in polynomial time,
        and $NP$ is the class of decision problems whose proposed solutions can
        be verified in polynomial time.
    \end{openproblem}

    \begin{remark}
        The usual slogan is: can every solution that can be checked quickly also
        be found quickly? The keywords are Cook, Karp, NP-completeness, and
        theoretical computer science.
    \end{remark}

    \begin{conjecture}[Birch and Swinnerton-Dyer Conjecture]
        For an elliptic curve $E$ over $\Q$, $\rk E(\Q)$ equals the order
        of vanishing of its $L$-function $L(E,s)$ at $s=1$.
    \end{conjecture}

    \begin{remark}
        The problem connects rational points on elliptic curves with analytic
        behavior of $L$-functions. The phrase ``$\rk$ equals order of vanishing''
        is the main quiz memory.
    \end{remark}

    \begin{conjecture}[Hodge Conjecture]
        On a smooth projective complex variety, every rational cohomology class
        of type $(p,p)$ should be a rational linear combination of classes of
        algebraic cycles.
    \end{conjecture}

    \begin{remark}
        This is one of the deepest bridges between topology and algebraic
        geometry. The short memory is
        \[
            \text{Hodge classes} \quad\longleftrightarrow\quad
            \text{algebraic cycles}.
        \]
    \end{remark}

    \begin{openproblem}[Navier--Stokes Existence and Smoothness]
        For the $3$-dimensional incompressible Navier--Stokes equations,
        determine whether smooth initial data always lead to global smooth
        solutions, or whether finite-time singularities can occur.
    \end{openproblem}

    \begin{remark}
        This is a PDE problem motivated by fluid mechanics. The quiz keywords
        are incompressible fluid, global regularity, singularity, and turbulence.
    \end{remark}

    \begin{openproblem}[Yang--Mills Existence and Mass Gap]
        Construct a mathematically rigorous quantum Yang--Mills theory on
        $\R^4$ for every compact simple gauge group and prove that it has a
        positive mass gap.
    \end{openproblem}

    \begin{remark}
        The problem comes from quantum field theory. In quiz form, the key words
        are gauge theory, quantum field theory, and mass gap.
    \end{remark}

    \section*{Classical Problems and Conjectures}

    The following problems are not all open problems. Some are solved theorems,
    but they remain famous because of their history, statements, or solution
    methods.

    \vspace{1em}

    {\small
    \renewcommand{\arraystretch}{1.18}
    \begin{longtable}{@{}L{0.23\textwidth}L{0.30\textwidth}L{0.20\textwidth}L{0.19\textwidth}@{}}
        \toprule
        Problem & Standard form & People and status & Quiz anchor \\
        \midrule
        \endfirsthead

        \toprule
        Problem & Standard form & People and status & Quiz anchor \\
        \midrule
        \endhead

        \midrule
        \multicolumn{4}{r}{Continued on the next page.} \\
        \endfoot

        \bottomrule
        \endlastfoot

        Fermat's Last Theorem
        & $x^n+y^n=z^n$ has no positive integer solutions for $n\ge 3$.
        & Fermat; proved by Wiles with Taylor--Wiles.
        & Elliptic curves and modular forms. \\

        Four Color Theorem
        & Every planar map can be colored with at most four colors.
        & Appel--Haken; later Robertson--Sanders--Seymour--Thomas.
        & First famous computer-assisted proof. \\

        Kepler Conjecture
        & The densest packing of congruent spheres in $\R^3$ is the face-centered
          cubic packing.
        & Kepler; proved by Hales.
        & Sphere packing; Flyspeck project. \\

        Thurston Geometrization Conjecture
        & Closed orientable $3$-manifolds decompose into pieces modeled on eight
          geometries.
        & Thurston; completed through Perelman's work.
        & Eight Thurston geometries. \\

        Continuum Hypothesis
        & There is no cardinality strictly between $|\N|$ and $|\R|$.
        & Cantor; independent of ZFC by Gödel and Cohen.
        & Forcing; independence. \\

        Hilbert's Problems
        & A list of $23$ problems proposed in 1900.
        & Hilbert; many solved, some open or independent.
        & Roadmap for twentieth-century mathematics. \\

        Goldbach Conjecture
        & Every even integer greater than $2$ is a sum of two primes.
        & Goldbach; open.
        & Additive number theory. \\

        Twin Prime Conjecture
        & There are infinitely many primes $p$ such that $p+2$ is also prime.
        & Open; bounded gaps by Zhang, Maynard, Tao, Polymath.
        & Prime gaps. \\

        Collatz Conjecture
        & Iterating $n\mapsto n/2$ for even $n$ and $n\mapsto 3n+1$ for odd $n$
          eventually reaches $1$.
        & Collatz; open.
        & Simple statement, hard dynamics. \\

        $abc$ Conjecture
        & If $a+b=c$ and $a,b,c$ are coprime, then $c$ is controlled by the
          radical of $abc$, up to $\varepsilon$.
        & Masser--Oesterlé; open in the standard sense.
        & Radical, Diophantine equations. \\
    \end{longtable}
    }

    \begin{theorem}[Fermat's Last Theorem]
        For every integer $n\ge 3$, there do not exist positive integers
        $x,y,z$ such that
        \[
            x^n+y^n=z^n.
        \]
    \end{theorem}

    \begin{remark}
        Fermat wrote that he had a marvelous proof, but no proof appeared in
        his margin. Andrew Wiles proved the theorem through the modularity of
        semistable elliptic curves. The fast association is
        \[
            \text{Fermat's Last Theorem}
            \quad\longleftrightarrow\quad
            \text{Wiles}
            \quad\longleftrightarrow\quad
            \text{elliptic curves and modular forms}.
        \]
    \end{remark}

    \begin{theorem}[Four Color Theorem]
        Every planar map can be colored with at most four colors so that any two
        adjacent regions have different colors.
    \end{theorem}

    \begin{remark}
        The theorem was proved by Kenneth Appel and Wolfgang Haken in 1976 with
        extensive computer assistance. This made it one of the most famous early
        examples of a computer-assisted proof.
    \end{remark}

    \begin{theorem}[Kepler Conjecture]
        The maximum density of a packing of congruent spheres in $\R^3$ is
        attained by the face-centered cubic packing and equals
        \[
            \frac{\pi}{3\sqrt{2}}.
        \]
    \end{theorem}

    \begin{remark}
        Thomas Hales proved the Kepler conjecture using a large computer-assisted
        argument. The Flyspeck project later produced a formal verification of
        the proof. In higher-dimensional sphere packing, Maryna Viazovska proved
        the optimality of the $E_8$ lattice in dimension $8$.
    \end{remark}

    \begin{theorem}[Thurston Geometrization Theorem]
        Every closed orientable $3$-manifold can be decomposed into pieces that
        admit one of the eight Thurston geometries:
        \[
            S^3,\ E^3,\ H^3,\ S^2\times\R,\ H^2\times\R,
            \ \widetilde{\SL}_2(\R),\ \mathrm{Nil},\ \mathrm{Sol}.
        \]
    \end{theorem}

    \begin{remark}
        The Poincaré conjecture is a special case of the geometrization program.
        The 2018 Science Quiz asked for the number of Thurston geometries; the
        answer is $8$.
    \end{remark}

    \begin{hypothesis}[Continuum Hypothesis]
        There is no set whose cardinality is strictly between the cardinality of
        the natural numbers and the cardinality of the real numbers. Equivalently,
        \[
            2^{\aleph_0}=\aleph_1.
        \]
    \end{hypothesis}

    \begin{remark}
        Kurt Gödel proved that CH cannot be disproved from ZFC if ZFC is
        consistent. Paul Cohen proved that CH cannot be proved from ZFC if ZFC
        is consistent. Cohen introduced forcing and received the Fields Medal in
        1966.
    \end{remark}

    \begin{fact}[Hilbert's Problems]
        David Hilbert presented $23$ problems at the 1900 International Congress
        of Mathematicians in Paris. They shaped much of twentieth-century
        mathematics. Useful anchors include:
        \begin{itemize}
            \item Hilbert's first problem $\leftrightarrow$ continuum hypothesis.
            \item Hilbert's second problem $\leftrightarrow$ consistency of arithmetic.
            \item Hilbert's tenth problem $\leftrightarrow$ Diophantine equations;
                  negative solution by Matiyasevich, building on Davis--Putnam--Robinson.
        \end{itemize}
    \end{fact}

    \begin{openproblem}[Goldbach Conjecture]
        Every even integer greater than $2$ is a sum of two prime numbers.
    \end{openproblem}

    \begin{remark}
        Goldbach is one of the easiest famous open problems to state. It belongs
        to additive number theory. The weak Goldbach conjecture, concerning sums
        of three odd primes, was proved by Harald Helfgott.
    \end{remark}

    \begin{openproblem}[Twin Prime Conjecture]
        There are infinitely many primes $p$ such that
        \[
            p+2
        \]
        is also prime.
    \end{openproblem}

    \begin{remark}
        The modern quiz keywords are bounded gaps between primes, Yitang Zhang,
        James Maynard, Terence Tao, and the Polymath project. These results do
        not prove the twin prime conjecture, but they prove that infinitely many
        prime pairs occur within some fixed bounded distance.
    \end{remark}

    \begin{openproblem}[Collatz Conjecture]
        Start with a positive integer $n$. If $n$ is even, replace it by $n/2$.
        If $n$ is odd, replace it by $3n+1$. The Collatz conjecture states that
        repeated iteration always eventually reaches $1$.
    \end{openproblem}

    \begin{remark}
        This problem is famous because the rule is elementary but the global
        behavior is still unknown. It is also called the $3n+1$ problem.
    \end{remark}

    \begin{conjecture}[$abc$ Conjecture]
        For every $\varepsilon>0$, there exists a constant $K_\varepsilon$ such
        that, whenever $a,b,c$ are pairwise coprime positive integers satisfying
        $a+b=c$, we have
        \[
            c<K_\varepsilon \rad(abc)^{1+\varepsilon},
        \]
        where $\rad(n)$ is the product of the distinct prime
        divisors of $n$.
    \end{conjecture}

    \begin{remark}
        The conjecture is powerful because it predicts that the additive
        relation $a+b=c$ strongly restricts the repeated prime factors of
        $abc$. The keyword is ``radical.''
    \end{remark}

    \section*{Modern Quiz-Relevant Conjecture Clusters}

    Some recent Science Quiz-style questions are built around a mathematician
    and the conjectures attached to that person's work. The most important
    example for current preparation is June Huh.

    \begin{fact}[June Huh and combinatorial log-concavity]
        June Huh's work connects combinatorics with algebraic geometry and Hodge
        theory. The most useful associations are:
        \begin{itemize}
            \item Read's conjecture $\leftrightarrow$ chromatic polynomials.
            \item Rota's conjecture $\leftrightarrow$ log-concavity for matroids.
            \item Mason's conjecture $\leftrightarrow$ independent sets of matroids.
            \item Hodge theory $\leftrightarrow$ the geometric method behind these
                  combinatorial log-concavity results.
        \end{itemize}
        In a quiz, if June Huh appears with a list of conjectures, the odd one
        out is often a famous conjecture from a different field, such as the
        Poincaré conjecture.
    \end{fact}
