    \chapter{Mathematics in 2026}

    This part is a working list of mathematical events, awards, and recent
    developments from the current Science Quiz season. It is meant to be updated
    repeatedly and replaced each year. Confirmed results are separated from topics
    that still require checking, so that speculation is never presented as fact.

    \section*{Confirmed Current Entries}

    \begin{itemize}
        \item \textbf{2025 Shaw Prize in Mathematical Sciences.} Kenji Fukaya was
              awarded the 2025 Shaw Prize in Mathematical Sciences for his work
              in symplectic geometry, the Fukaya category, symplectic topology,
              mirror symmetry, and gauge theory.

        \item \textbf{2025 AI and the IMO.} Systems reported by Google DeepMind
              and OpenAI reached gold-medal-level scores on the 2025 International
              Mathematical Olympiad problems. The useful Science Quiz distinction
              is that Google DeepMind's performance was evaluated through the IMO
              process, whereas OpenAI reported an independent evaluation. Quiz
              keywords: mathematical reasoning, olympiad problems, natural-language
              proofs, formal verification, and AI-assisted mathematics.

        \item \textbf{2026 ICM and IMU General Assembly.} The International
              Congress of Mathematicians is scheduled for Philadelphia, USA, from
              July 23 to July 30, 2026. The 20th IMU General Assembly is scheduled
              for New York City from July 20 to July 21, 2026.

        \item \textbf{2026 IMU prizes at ICM.} The ICM opening ceremony is the
              key moment to record the Fields Medals, the IMU Abacus Medal, the
              Carl Friedrich Gauss Prize, and the Chern Medal Award. The Leelavati
              Prize is associated with the closing ceremony. The laureates should
              be added only after the official IMU announcement.

        \item \textbf{2026 Abel Prize.} Gerd Faltings received the 2026 Abel
              Prize for introducing powerful tools in arithmetic geometry and
              resolving long-standing Diophantine conjectures of Mordell and Lang.
              Quiz keywords: arithmetic geometry, Mordell conjecture, Faltings'
              theorem, Mordell--Lang conjecture, and Diophantine geometry.

        \item \textbf{2026 Breakthrough Prize in Mathematics.} Frank Merle
              received the 2026 Breakthrough Prize in Mathematics for work on
              nonlinear evolution equations, including stability, singularity
              formation, and resolution into solitons.

        \item \textbf{2026 Clay Research Awards.} The awards recognized three
              groups: Tuomas Orponen, Pablo Shmerkin, Hong Wang, and Joshua Zahl;
              Robert Burklund, Jeremy Hahn, Ishan Levy, and Tomer Schlank; and Yu
              Deng and Zaher Hani. Quiz keywords: the Furstenberg set conjecture,
              the Kakeya conjecture, the Telescope Conjecture, and the Boltzmann
              equation.

        \item \textbf{Moving sofa problem.} Jineon Baek posted the preprint
              \textit{Optimality of Gerver's Sofa}, claiming a resolution of the
              moving sofa problem by proving that Gerver's sofa is optimal. This
              belongs on the Korea-related watchlist, but its publication and
              independent verification status should be checked before the final
              printing.
    \end{itemize}

    \section*{Topics to Develop and Verify}

    \begin{itemize}
        \item \textbf{ICM 2026 prize results.} Add the 2026 Fields Medalists,
              IMU Abacus Medalist, Gauss Prize laureate, Chern Medal laureate, and
              Leelavati Prize laureate immediately after the official ceremonies.
              Record each laureate together with one or two reliable mathematical
              keywords rather than candidate speculation.

        \item \textbf{ICM 2026 mathematical themes.} Review the plenary and
              invited lectures after the congress and identify a small number of
              fields, theorems, or research programs that are likely to become
              Science Quiz questions.

        \item \textbf{IMO 2026.} After the official record is complete, add the
              host city, participating countries, team results, medal cutoffs,
              perfect scores, and especially memorable solutions. Separate human
              competition results from AI evaluations.

        \item \textbf{AI theorem proving and formalization.} Track the increasing
              use of Lean and other proof assistants, autoformalization, verified
              proof search, and human--AI collaboration. Distinguish clearly among
              olympiad benchmark performance, formal machine-checked proofs, and
              genuine progress on research-level open problems.

        \item \textbf{Geometric Langlands program.} Add the 2025 Breakthrough
              Prize connection to Dennis Gaitsgory and the large collaborative
              proof of the geometric Langlands conjecture only after checking the
              precise mathematical scope and publication status.

        \item \textbf{2026 Shaw Prize and Wolf Prize.} Check the official Shaw
              Prize Foundation and Wolf Foundation announcements. Do not infer a
              laureate or even a mathematics award from an informal prediction.

        \item \textbf{Korea-related mathematics news.} Check major prizes,
              appointments, deaths, young mathematicians, KIAS and IBS activity,
              and work connected to \,\POSTECH, \,\KAIST, and Professor June Huh.
              Give priority to items with a clear theorem, problem, or named object
              that can become a short-answer question.

        \item \textbf{Open-problem progress.} Monitor bounded gaps between
              primes, the Navier--Stokes problem, geometric analysis, arithmetic
              geometry, and other major research areas. Include an item only when
              the exact theorem and its status can be stated from a primary source;
              numerical evidence, a preprint claim, and a completed proof must not
              be conflated.

        \item \textbf{2026 \POSTECH--\KAIST{} Science War.} Add season-specific
              trends after practice problems, team notes, and actual match
              information become available.
    \end{itemize}

    \begin{remark}
        Sensational claims that a named AI system has solved a famous decades-old
        conjecture are not included without a public proof, identifiable authors,
        and independent mathematical verification. This chapter should remain a
        reliable quiz checklist rather than a collection of rumors.
    \end{remark}
