    \begin{problem}{2010.1}
        Determine whether there exist positive integers $x$ and $y$ satisfying
        \[
            x^2+y^2=963.
        \]
    \end{problem}

    \begin{solution}
        No such positive integers exist. For any integer $a$, we have
        \[
            a^2\equiv 0 \text{ or } 1\pmod{4}.
        \]
        Hence the sum of two integer squares can be congruent only to
        \[
            0,\quad 1,\quad \text{or}\quad 2\pmod{4}.
        \]
        However,
        \[
            963\equiv 3\pmod{4}.
        \]
        Therefore $963$ cannot be written as the sum of two integer squares.

        Therefore, the answer is $\boxed{\text{no}}$.
    \end{solution}

    \begin{problem}{2010.2}
        Identify the mathematician who became famous for refusing the Fields Medal after solving the Poincar\'e conjecture.
    \end{problem}

    \begin{solution}
        Grigori Perelman proved the Poincar\'e conjecture using Ricci flow. He was awarded the Fields Medal in 2006, but declined it.

        Therefore, the answer is $\boxed{\text{Grigori Perelman}}$.
    \end{solution}

    \begin{problem}{2010.3}
        Determine the country of origin of John Charles Fields, after whom the Fields Medal is named.
    \end{problem}

    \begin{solution}
        John Charles Fields was a Canadian mathematician. The Fields Medal is named after him.

        Therefore, the answer is $\boxed{\text{Canada}}$.
    \end{solution}

    \begin{problem}{2010.4}
        Determine the maximum possible number of distinct eigenvalues of a $5\times 5$ matrix.
    \end{problem}

    \begin{solution}
        The eigenvalues of a $5\times 5$ matrix $\mathbf{A}$ are the roots of its characteristic polynomial
        \[
            p_{\mathbf{A}}(\lambda)=\det(\lambda\mathbf{I}_5-\mathbf{A}).
        \]
        This polynomial has degree $5$. Therefore it can have at most $5$ distinct roots, and hence $\mathbf{A}$ can have at most $5$ distinct eigenvalues.

        Therefore, the answer is $\boxed{5}$.
    \end{solution}

    \begin{problem}{2010.5}
        In 1742, a Prussian mathematician wrote to Euler and proposed a conjecture. Euler later stated it in the following form:
        \[
            \text{Every even integer greater than }2\text{ can be expressed as the sum of two primes.}
        \]
        This is an old unsolved problem in number theory. It has been verified for very large numbers, but it has not been proved in general. Identify this conjecture.
    \end{problem}

    \begin{solution}
        The statement that every even integer greater than $2$ can be expressed as the sum of two primes is known as Goldbach's conjecture.

        Therefore, the answer is $\boxed{\text{Goldbach's conjecture}}$.
    \end{solution}

    \begin{problem}{2010.6}
        Brownian motion is mathematically modeled by a random process named after the American mathematician who developed its mathematical model. This mathematician also founded cybernetics.

        The process has increments that are independent Gaussian random variables with mean $0$. The mathematician studied real analysis, harmonic analysis, series, and probability theory. Identify this mathematician.
    \end{problem}

    \begin{solution}
        The random process described in the problem is the Wiener process, the standard mathematical model of Brownian motion. It is named after Norbert Wiener.

        Therefore, the answer is $\boxed{\text{Norbert Wiener}}$.
    \end{solution}

    \begin{problem}{2010.7}
        This set is a closed uncountable set and is also perfect. Although it is uncountable, it has Lebesgue measure $0$. It is self-similar and is often used in the construction of fractals. It is obtained from the closed interval $[0,1]$ by repeatedly removing the open middle third, then doing the same to the remaining intervals. It is named after the mathematician widely known as the founder of set theory. Identify this set.
    \end{problem}

    \begin{solution}
        The set obtained by repeatedly removing the open middle third from $[0,1]$ is the Cantor set. It is closed, perfect, uncountable, self-similar, and has Lebesgue measure $0$.

        Therefore, the answer is $\boxed{\text{Cantor set}}$.
    \end{solution}

    \begin{problem}{2010.8}
        Determine the year in which Terence Tao received the Fields Medal.
    \end{problem}

    \begin{solution}
        Terence Tao received the Fields Medal at the 2006 International Congress of Mathematicians.

        Therefore, the answer is $\boxed{2006}$.
    \end{solution}

    \begin{problem}{2011.1}
        Identify the international mathematical congress, held in South Korea in 2014, at which the Fields Medal is awarded.
    \end{problem}

    \begin{solution}
        The Fields Medal is awarded at the International Congress of Mathematicians, commonly abbreviated as ICM. The 2014 ICM was held in Seoul, South Korea.

        Therefore, the answer is $\boxed{\text{International Congress of Mathematicians, or ICM}}$.
    \end{solution}

    \begin{problem}{2011.2}
        Let $\mathbf{A}$ and $\mathbf{B}$ be $n\times n$ matrices. If
        \[
            \mathbf{A}\mathbf{B}-\mathbf{I}=\mathbf{0},
        \]
        where $\mathbf{I}$ is the identity matrix. Determine
        \[
            \mathbf{B}\mathbf{A}-\mathbf{I}.
        \]
    \end{problem}

    \begin{solution}
        The condition
        \[
            \mathbf{A}\mathbf{B}-\mathbf{I}=\mathbf{0}
        \]
        means that
        \[
            \mathbf{A}\mathbf{B}=\mathbf{I}.
        \]
        Since $\mathbf{A}$ and $\mathbf{B}$ are square matrices, this implies that $\mathbf{B}$ is the inverse of $\mathbf{A}$. Hence
        \[
            \mathbf{B}\mathbf{A}=\mathbf{I}.
        \]
        Therefore,
        \[
            \mathbf{B}\mathbf{A}-\mathbf{I}
            =
            \mathbf{I}-\mathbf{I}
            =
            \mathbf{0}.
        \]

        Therefore, the answer is $\boxed{\mathbf{0}}$.
    \end{solution}

    \begin{problem}{2011.3}
        Differentiate $(1+x)^{10}$ three times, substitute $x=0$, and determine the resulting value.
    \end{problem}

    \begin{solution}
        We compute the third derivative:
        \[
            \frac{\dd^3}{\dd x^3}(1+x)^{10}
            =
            10\cdot 9\cdot 8(1+x)^7.
        \]
        Substituting $x=0$, we get
        \[
            10\cdot 9\cdot 8(1+0)^7
            =
            720.
        \]
        Therefore, the answer is $\boxed{720}$.
    \end{solution}

    \begin{problem}{2011.4}
        Give at least two real-valued functions defined on $\R$ that remain unchanged after being differentiated twice. In other words, give at least two functions $f:\R\to\R$ satisfying
        \[
            f''=f.
        \]
    \end{problem}

    \begin{solution}
        Two simple examples are
        \[
            f(x)=e^x
            \qquad\text{and}\qquad
            g(x)=e^{-x}.
        \]
        Indeed,
        \[
            \frac{\dd^2}{\dd x^2}e^x=e^x
        \]
        and
        \[
            \frac{\dd^2}{\dd x^2}e^{-x}=e^{-x}.
        \]
        Therefore, both functions remain unchanged after being differentiated twice.

        Therefore, two valid examples are $\boxed{e^x\text{ and }e^{-x}}$.
    \end{solution}

    \begin{problem}{2011.5}
        Let $\mathbf{A}$ be a $5\times 5$ matrix such that every entry in the $i$-th row is equal to $i$. Find the sum of the five eigenvalues of $\mathbf{A}$, counted with multiplicity.
    \end{problem}

    \begin{solution}
        The sum of the eigenvalues of a square matrix, counted with algebraic multiplicity, is equal to $\tr(\mathbf{A})$.

        In this matrix, the diagonal entries are
        \[
            1,\quad 2,\quad 3,\quad 4,\quad 5.
        \]
        Therefore,
        \[
            \tr(\mathbf{A})
            =
            1+2+3+4+5
            =
            15.
        \]
        Hence the sum of the five eigenvalues of $\mathbf{A}$, counted with multiplicity, is $\boxed{15}$.
    \end{solution}

    \begin{problem}{2011.6}
        A train takes $22$ seconds to completely cross an $82\mathrm{m}$ bridge. When its speed is doubled, it takes $50$ seconds to completely cross a $706\mathrm{m}$ bridge. Determine the length of the train.
    \end{problem}

    \begin{solution}
        Let $L$ be the length of the train, and let $v$ be its original speed in meters per second.

        To completely cross a bridge, the train must travel the length of the bridge plus its own length. Hence
        \[
            \frac{L+82}{v}=22.
        \]
        When the speed is doubled, the train takes $50$ seconds to completely cross a $706\mathrm{m}$ bridge, so
        \[
            \frac{L+706}{2v}=50.
        \]
        Equivalently,
        \[
            L+82=22v,
            \qquad
            L+706=100v.
        \]
        Subtracting the first equation from the second gives
        \[
            624=78v,
        \]
        so
        \[
            v=8.
        \]
        Therefore,
        \[
            L+82=22\cdot 8=176,
        \]
        and hence
        \[
            L=94.
        \]

        Thus the length of the train is $\boxed{94\mathrm{m}}$.
    \end{solution}

    \begin{problem}{2011.7}
        Form an integer using the digits
        \[
            1,2,3,4,6,7,8,9.
        \]
        Each of these digits must be used at least once, and digits may be repeated. Determine the smallest such integer divisible by all of
        \[
            1,2,3,4,6,7,8,9.
        \]
    \end{problem}

    \begin{solution}
        The number must be divisible by
        \[
            \lcm(1,2,3,4,6,7,8,9)=504.
        \]
        Thus it must be divisible by $7$, $8$, and $9$.

        If each digit is used exactly once, then the sum of the digits is
        \[
            1+2+3+4+6+7+8+9=40,
        \]
        which is not divisible by $9$. Hence no $8$-digit number works.

        If we use exactly one extra digit $d$, then the digit sum becomes $40+d$. For divisibility by $9$, we need
        \[
            40+d\equiv 0\pmod{9},
        \]
        so
        \[
            d\equiv 5\pmod{9}.
        \]
        But the digit $5$ is not allowed. Hence no $9$-digit number works.

        Therefore, the answer must have at least $10$ digits. For a $10$-digit number, we need to add two extra digits. Their sum must be congruent to $5$ modulo $9$. Among the allowed digits, the possible extra pairs are
        \[
            (1,4),\qquad (2,3),\qquad (6,8),\qquad (7,7).
        \]
        To make the number as small as possible, we should try the lexicographically smallest multiset first, namely
        \[
            1,1,2,3,4,4,6,7,8,9.
        \]

        With these digits, the smallest possible beginning is
        \[
            112344.
        \]
        It remains to arrange the last four digits $6,7,8,9$ so that the whole number is divisible by $504$.

        Since the digit sum is already divisible by $9$, it remains to check divisibility by $7$ and $8$. The last three digits must be divisible by $8$. Among the permutations of three of the digits $6,7,8,9$, the possible last three digits divisible by $8$ are
        \[
            768,\qquad 896,\qquad 968,\qquad 976.
        \]
        Hence the only candidates with prefix $112344$ are
        \[
            1123449768,\qquad
            1123447896,\qquad
            1123447968,\qquad
            1123448976.
        \]
        Checking divisibility by $7$, only
        \[
            1123449768
        \]
        is divisible by $7$. Indeed,
        \[
            1123449768=504\cdot 2229067.
        \]
        Therefore it is divisible by all of
        \[
            1,2,3,4,6,7,8,9.
        \]

        Thus the smallest possible integer is $\boxed{1123449768}$.
    \end{solution}

    \begin{problem}{2011.8}
        Determine the greatest integer less than or equal to
        \[
            -\frac{1}{\ln\left(\frac{365}{366}\right)}.
        \]

        \begin{hint}
            Use the Taylor approximation of $\ln(1-x)$.
        \end{hint}
    \end{problem}

    \begin{solution}
        We write
        \[
            \frac{365}{366}=1-\frac{1}{366}.
        \]
        Using
        \[
            \ln(1-x)
            =
            -x-\frac{x^2}{2}-\frac{x^3}{3}-\cdots,
        \]
        with $x=\dfrac{1}{366}$, we get
        \[
            -\ln\left(\frac{365}{366}\right)
            =
            \frac{1}{366}
            +
            \frac{1}{2\cdot 366^2}
            +
            \frac{1}{3\cdot 366^3}
            +
            \cdots.
        \]
        Hence
        \[
            \frac{1}{366}
            <
            -\ln\left(\frac{365}{366}\right)
            <
            \frac{1}{365}.
        \]
        Taking reciprocals gives
        \[
            365
            <
            -\frac{1}{\ln\left(\frac{365}{366}\right)}
            <
            366.
        \]
        Therefore, the greatest integer less than or equal to the given number is $\boxed{365}$.
    \end{solution}

    \begin{problem}{2011.9}
        Evaluate the integral
        \[
            \int_0^{\pi/2}
            \frac{1}{1+\tan^{\sqrt{2}}t}\dd t.
        \]

        \begin{hint}
            Use the substitution $x=\dfrac{\pi}{2}-t$.
        \end{hint}
    \end{problem}

    \begin{solution}
        Let
        \[
            I
            =
            \int_0^{\pi/2}
            \frac{1}{1+\tan^{\sqrt{2}}t}\dd t.
        \]
        Use the substitution
        \[
            x=\frac{\pi}{2}-t.
        \]
        Then
        \[
            \tan t
            =
            \tan\left(\frac{\pi}{2}-x\right)
            =
            \cot x
            =
            \frac{1}{\tan x}.
        \]
        Hence
        \[
            I
            =
            \int_0^{\pi/2}
            \frac{1}{1+\cot^{\sqrt{2}}x}\dd x
            =
            \int_0^{\pi/2}
            \frac{\tan^{\sqrt{2}}x}{1+\tan^{\sqrt{2}}x}\dd x.
        \]
        Adding this expression to the original expression for $I$, we obtain
        \[
            2I
            =
            \int_0^{\pi/2}
            \left(
                \frac{1}{1+\tan^{\sqrt{2}}x}
                +
                \frac{\tan^{\sqrt{2}}x}{1+\tan^{\sqrt{2}}x}
            \right)\dd x
            =
            \int_0^{\pi/2}1\dd x
            =
            \frac{\pi}{2}.
        \]
        Therefore,
        \[
            I=\frac{\pi}{4}.
        \]
        Therefore, the answer is $\boxed{\dfrac{\pi}{4}}$.
    \end{solution}

    \begin{problem}{2011.10}
        This French mathematician was born on December 8, 1865, and died on October 17, 1963, living to the age of $97$. He is well known for his proof of the prime number theorem, which states that the number of primes less than or equal to $x$ is asymptotically approximated by $\dfrac{x}{\ln x}$ for large $x$. He was also personally connected to the Dreyfus affair and became a strong supporter of Zionism.

        Identify this mathematician.

        \begin{hint}
            The following clues are useful:
            \begin{enumerate}
                \item[(a)] The Cauchy--\underline{\hspace{2cm}} theorem gives the radius of convergence of a power series.
                \item[(b)] A \underline{\hspace{2cm}} matrix is a square matrix whose entries are all $+1$ or $-1$, and such matrices are also used in error-correcting codes.
            \end{enumerate}
        \end{hint}
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is Jacques Hadamard.

        The first hint refers to the Cauchy--Hadamard theorem, which gives the radius of convergence of a power series. The second hint refers to a Hadamard matrix, a square matrix whose entries are all $+1$ or $-1$ and whose rows are mutually orthogonal.

        Hadamard is also famous for proving the prime number theorem in 1896, independently of Charles Jean de la Vallée Poussin. Therefore, the answer is $\boxed{\text{Jacques Hadamard}}$.
    \end{solution}

    \begin{problem}{2012.1}
        This British mathematical genius died of cyanide poisoning shortly before his $42$nd birthday. A famous account says that he died after eating an apple laced with cyanide. Identify this mathematician.
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is Alan Turing.

        Turing was a British mathematician, logician, and computer scientist. He made fundamental contributions to computability theory through the concept of the Turing machine, and he also played a crucial role in codebreaking during World War II.

        Turing died in 1954 from cyanide poisoning, shortly before his $42$nd birthday. The story involving an apple laced with cyanide is widely known, although some details of the event remain historically uncertain.

        Therefore, the answer is $\boxed{\text{Alan Turing}}$.
    \end{solution}

    \begin{problem}{2012.2}
        Evaluate the infinite series
        \[
            1+\frac{1}{2^2}+\frac{1}{3^2}+\cdots.
        \]
    \end{problem}

    \begin{solution}
        The given series is
        \[
            \sum_{n=1}^{\infty}\frac{1}{n^2}.
        \]
        This is the famous Basel problem. Euler proved that
        \[
            \sum_{n=1}^{\infty}\frac{1}{n^2}
            =
            \frac{\pi^2}{6}.
        \]
        Therefore, the value of the series is $\boxed{\dfrac{\pi^2}{6}}$.
    \end{solution}

    \begin{problem}{2012.3}
        Identify the mathematician who first proved the law of quadratic reciprocity and discovered more than seven proofs of it.
    \end{problem}

    \begin{solution}
        The law of quadratic reciprocity is one of the central theorems in elementary number theory. It describes a deep symmetry between the solvability of
        \[
            x^2\equiv p \pmod q
            \qquad\text{and}\qquad
            x^2\equiv q \pmod p
        \]
        for odd primes $p$ and $q$.

        Carl Friedrich Gauss gave the first complete proof of the law of quadratic reciprocity. He regarded it as one of the jewels of number theory and found several different proofs of the theorem.

        Therefore, the answer is $\boxed{\text{Carl Friedrich Gauss}}$.
    \end{solution}

    \begin{problem}{2012.4}
        Let $\mathbf{A}$ be a $10\times 10$ matrix such that every entry in the $k$-th row is equal to $k$. Find the sum of the eigenvalues of $\mathbf{A}$, counted with multiplicity.
    \end{problem}

    \begin{solution}
        The sum of the eigenvalues of a square matrix, counted with algebraic multiplicity, is equal to $\tr(\mathbf{A})$.

        In this matrix, the diagonal entries are
        \[
            1,\quad 2,\quad 3,\quad \dots,\quad 10.
        \]
        Therefore,
        \[
            \tr(\mathbf{A})
            =
            1+2+3+\cdots+10
            =
            \frac{10\cdot 11}{2}
            =
            55.
        \]
        Hence the sum of the eigenvalues of $\mathbf{A}$, counted with multiplicity, is $\boxed{55}$.
    \end{solution}

    \begin{problem}{2012.5}
        This philosopher was born as the youngest son of a wealthy Austrian steel magnate who supported many artists. He showed talent in several fields, especially mathematics and philosophy. His work had a major influence on logical positivism and on the philosophy of language, and it also influenced the humanities, social sciences, and the arts. His later philosophy is especially known for emphasizing ordinary language use.

        Identify this person.
    \end{problem}

    \begin{solution}
        The person described in the problem is Ludwig Wittgenstein.

        Wittgenstein was born into a wealthy Austrian family and studied logic, mathematics, and philosophy. His early work, especially the \textit{Tractatus Logico-Philosophicus}, strongly influenced logical positivism and the Vienna Circle.

        Later, in works such as \textit{Philosophical Investigations}, Wittgenstein emphasized the importance of ordinary language and argued that the meaning of a word is closely connected to its use in language.

        Therefore, the answer is $\boxed{\text{Ludwig Wittgenstein}}$.
    \end{solution}

    \begin{problem}{2012.6}
        The verified original statement and any accompanying figure are unavailable. This entry is retained and will be completed when the source is recovered.
    \end{problem}

    \begin{solution}
        The original source is unavailable, so the problem and its solution cannot yet be reconstructed reliably. This entry is retained to preserve the original numbering.
    \end{solution}

    \begin{problem}{2014.1}
        This twentieth-century German mathematician is well known for a theorem stating that symmetries of Hamiltonian dynamical systems correspond to conservation laws. Identify this mathematician.
    \end{problem}

    \begin{solution}
        The theorem described in the problem is Noether's theorem. It states, roughly speaking, that continuous symmetries of a physical system correspond to conserved quantities. For example, time-translation symmetry corresponds to conservation of energy, and spatial-translation symmetry corresponds to conservation of momentum.

        The mathematician associated with this theorem is Emmy Noether, one of the most important algebraists of the twentieth century.

        Therefore, the answer is $\boxed{\text{Emmy Noether}}$.
    \end{solution}

    \begin{problem}{2014.2}
        In the famous Chinese mathematical classic \textit{The Nine Chapters on the Mathematical Art}, often compared with Euclid's \textit{Elements} as a foundational mathematical classic, the value of $\pi$ appears as $3$. Later, Liu Hui, who wrote a commentary on \textit{The Nine Chapters}, obtained the approximation $3.14$ by refining the method of inscribed polygons. Identify the ancient Chinese mathematician who further refined this method and obtained the approximations
        \[
            \frac{355}{113}
            \qquad\text{and}\qquad
            \frac{22}{7}
        \]
        for $\pi$.
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is Zu Chongzhi.

        Liu Hui improved the approximation of $\pi$ by using polygons with many sides. Zu Chongzhi later obtained the very accurate bounds
        \[
            3.1415926<\pi<3.1415927,
        \]
        and gave two famous rational approximations:
        \[
            \frac{22}{7}
            \qquad\text{and}\qquad
            \frac{355}{113}.
        \]
        The approximation $\dfrac{355}{113}$ is especially accurate and is traditionally known as the Milü, or ``accurate ratio.''

        Therefore, the answer is $\boxed{\text{Zu Chongzhi}}$.
    \end{solution}

    \begin{problem}{2014.3}
        Consider the $2014\times 2014$ matrix whose diagonal entries are all $0$ and whose off-diagonal entries are all $1$. Find all eigenvalues of this matrix.
    \end{problem}

    \begin{solution}
        Let $\mathbf{A}$ be the given matrix, and let $\mathbf{J}$ be the $2014\times 2014$ matrix whose entries are all $1$. Then
        \[
            \mathbf{A}=\mathbf{J}-\mathbf{I}.
        \]

        The matrix $\mathbf{J}$ has eigenvalue $2014$ for the vector
        \[
            \mathbf{1}=(1,1,\dots,1)^{\mathsf T},
        \]
        since
        \[
            \mathbf{J}\mathbf{1}=2014\mathbf{1}.
        \]
        Therefore,
        \[
            \mathbf{A}\mathbf{1}
            =
            (\mathbf{J}-\mathbf{I})\mathbf{1}
            =
            2014\mathbf{1}-\mathbf{1}
            =
            2013\mathbf{1}.
        \]
        Hence $2013$ is an eigenvalue of $\mathbf{A}$.

        Now take any vector $\mathbf{v}=(v_1,\dots,v_{2014})^{\mathsf T}$ such that
        \[
            v_1+\cdots+v_{2014}=0.
        \]
        Then $\mathbf{J}\mathbf{v}=\mathbf{0}$, so
        \[
            \mathbf{A}\mathbf{v}
            =
            (\mathbf{J}-\mathbf{I})\mathbf{v}
            =
            -\mathbf{v}.
        \]
        Thus $-1$ is an eigenvalue. The subspace of vectors whose coordinates sum to $0$ has dimension $2013$, so the multiplicity of $-1$ is $2013$.

        Therefore, the eigenvalues are
        \[
            \boxed{2013 \text{ with multiplicity } 1,\qquad -1 \text{ with multiplicity } 2013}.
        \]
    \end{solution}

    \begin{problem}{2014.4}
        Let $m\geq 3$ and $n\geq 3$. Suppose that
        \[
            \{\mathbf{v}_1,\mathbf{v}_2,\mathbf{v}_3\}\subseteq \R^m,
            \qquad
            \{\mathbf{w}_1,\mathbf{w}_2,\mathbf{w}_3\}\subseteq \R^n.
        \]
        Assume that
        \[
            V=\Span\{\mathbf{v}_1,\mathbf{v}_2,\mathbf{v}_3\}
        \]
        is a $2$-dimensional subspace of $\R^m$, and that
        \[
            W=\Span\{\mathbf{w}_1,\mathbf{w}_2,\mathbf{w}_3\}
        \]
        is a $3$-dimensional subspace of $\R^n$. Determine $\rk(\mathbf{A})$ for the $m\times n$ matrix
        \[
            \mathbf{A}
            =
            \mathbf{v}_1\mathbf{w}_1^{\mathsf T}
            +
            \mathbf{v}_2\mathbf{w}_2^{\mathsf T}
            +
            \mathbf{v}_3\mathbf{w}_3^{\mathsf T}.
        \]
    \end{problem}

    \begin{solution}
        Let
        \[
            \mathbf{P}
            =
            \begin{pmatrix}
                \mathbf{v}_1 & \mathbf{v}_2 & \mathbf{v}_3
            \end{pmatrix},
            \qquad
            \mathbf{Q}
            =
            \begin{pmatrix}
                \mathbf{w}_1 & \mathbf{w}_2 & \mathbf{w}_3
            \end{pmatrix}.
        \]
        Then
        \[
            \mathbf{A}
            =
            \mathbf{P}\mathbf{Q}^{\mathsf T}.
        \]

        Since $\mathbf{v}_1,\mathbf{v}_2,\mathbf{v}_3$ span a $2$-dimensional subspace, we have
        \[
            \rk(\mathbf{P})=2.
        \]
        Since $\mathbf{w}_1,\mathbf{w}_2,\mathbf{w}_3$ span a $3$-dimensional subspace, they are linearly independent, and hence
        \[
            \rk(\mathbf{Q})=3.
        \]
        Therefore, the linear map $\mathbf{Q}^{\mathsf T}:\R^n\to\R^3$ is surjective.

        It follows that the image of $\mathbf{A}=\mathbf{P}\mathbf{Q}^{\mathsf T}$ is
        \[
            \image(\mathbf{A})
            =
            \mathbf{P}(\image(\mathbf{Q}^{\mathsf T}))
            =
            \mathbf{P}(\R^3)
            =
            \Span\{\mathbf{v}_1,\mathbf{v}_2,\mathbf{v}_3\}.
        \]
        This space has dimension $2$. Hence
        \[
            \rk(\mathbf{A})=2.
        \]

        Therefore, the answer is $\boxed{2}$.
    \end{solution}

    \begin{problem}{2015.1}
        Read the following passage and fill in the blank with the title of the mathematical classic, either in English or in Korean.

        \begin{quote}
            For $2000$ years, \underline{\hspace{2cm}} was not only the core of mathematical education but also at the heart of Western culture. The most glowing tributes to \underline{\hspace{2cm}} do not, in fact, come from mathematicians, but from philosophers, politicians, and others.

            Here are some examples:
            \begin{enumerate}
                \item[(a)] He studied and nearly mastered the six books of \underline{\hspace{2cm}} since he was a member of Congress. He regrets his want of education, and does what he can to supply the want.

                \hfill Abraham Lincoln, \textit{Short Autobiography}

                \item[(b)] He studied \underline{\hspace{2cm}} until he could demonstrate with ease all the propositions in the six books.

                \hfill \textit{Herndon's Life of Lincoln}

                \item[(c)] At the age of eleven, I began \underline{\hspace{2cm}}. This was one of the great events of my life, as dazzling as first love. I had not imagined there was anything so delicious in the world.

                \hfill Bertrand Russell, \textit{Autobiography}, Vol. 1
            \end{enumerate}
        \end{quote}
    \end{problem}

    \begin{solution}
        The passage is referring to Euclid's \textit{Elements}, one of the most influential works in the history of mathematics.

        For centuries, Euclid's \textit{Elements} served as a standard model of rigorous mathematical reasoning and was central to mathematical education. The quotations from Abraham Lincoln and Bertrand Russell emphasize how deeply the book influenced not only mathematicians, but also philosophers, politicians, and intellectual culture more broadly.

        Therefore, the answer is $\boxed{\text{\textit{The Elements}},\ \text{or Euclid's \textit{Elements}}}$.
    \end{solution}

    \begin{problem}{2015.2}
        Let $\mathbf{A}$ be an invertible square matrix. Suppose that, for a suitable permutation matrix $\mathbf{P}$, Gaussian elimination gives an LU-decomposition
        \[
            \mathbf{P}\mathbf{A}=\mathbf{L}\mathbf{U}.
        \]
        It may also happen that another permutation matrix $\widetilde{\mathbf{P}}$ gives another LU-decomposition
        \[
            \widetilde{\mathbf{P}}\mathbf{A}
            =
            \widetilde{\mathbf{L}}\widetilde{\mathbf{U}}.
        \]
        Show that the sets of absolute values of the pivots can be different. In other words, show by example that
        \[
            \{|u_{kk}|:1\leq k\leq n\}
            \neq
            \{|\widetilde{u}_{kk}|:1\leq k\leq n\}.
        \]

        \begin{hint}
            Look for a counterexample in the $2\times2$ case.
        \end{hint}
    \end{problem}

    \begin{solution}
        Consider the invertible matrix
        \[
            \mathbf{A}
            =
            \begin{pmatrix}
                1 & \frac{1}{2} \\
                \frac{1}{3} & 1
            \end{pmatrix}.
        \]
        With no row exchange, we have the LU-decomposition
        \[
            \mathbf{A}
            =
            \begin{pmatrix}
                1 & 0 \\
                \frac{1}{3} & 1
            \end{pmatrix}
            \begin{pmatrix}
                1 & \frac{1}{2} \\
                0 & \frac{5}{6}
            \end{pmatrix}.
        \]
        Hence the absolute values of the pivots are
        \[
            \left\{1,\frac{5}{6}\right\}.
        \]

        Now let $\widetilde{\mathbf{P}}$ be the permutation matrix that swaps the two rows. Then
        \[
            \widetilde{\mathbf{P}}\mathbf{A}
            =
            \begin{pmatrix}
                \frac{1}{3} & 1 \\
                1 & \frac{1}{2}
            \end{pmatrix}.
        \]
        This matrix has the LU-decomposition
        \[
            \widetilde{\mathbf{P}}\mathbf{A}
            =
            \begin{pmatrix}
                1 & 0 \\
                3 & 1
            \end{pmatrix}
            \begin{pmatrix}
                \frac{1}{3} & 1 \\
                0 & -\frac{5}{2}
            \end{pmatrix}.
        \]
        Hence the absolute values of the pivots are
        \[
            \left\{\frac{1}{3},\frac{5}{2}\right\}.
        \]

        Therefore,
        \[
            \left\{1,\frac{5}{6}\right\}
            \neq
            \left\{\frac{1}{3},\frac{5}{2}\right\}.
        \]
        This shows that the set of absolute values of the pivots can depend on the chosen row permutation.
    \end{solution}

    \begin{problem}{2015.3}
        In $\R^2$, the so-called Mercedes-Benz system is defined by
        \[
            \{\mathbf{v}_0,\mathbf{v}_1,\mathbf{v}_2\},
            \qquad
            \mathbf{v}_k
            =
            \begin{pmatrix}
                \cos\left(\frac{2k\pi}{3}\right) \\
                \sin\left(\frac{2k\pi}{3}\right)
            \end{pmatrix}
            \quad
            (k=0,1,2).
        \]
        Find vectors $\mathbf{w}_0,\mathbf{w}_1,\mathbf{w}_2\in\R^3$ such that they have the same length, are mutually orthogonal, have positive $z$-coordinates, and satisfy
        \[
            \mathbf{v}_k=P_{xy}(\mathbf{w}_k)
            \qquad
            (k=0,1,2),
        \]
        where $P_{xy}$ denotes the projection onto the $xy$-plane. Here $\R^2$ is regarded as the subspace of $\R^3$ with $z$-coordinate equal to $0$.
    \end{problem}

    \begin{solution}
        Since $P_{xy}(\mathbf{w}_k)=\mathbf{v}_k$, each $\mathbf{w}_k$ must have the form
        \[
            \mathbf{w}_k
            =
            \begin{pmatrix}
                \cos\left(\frac{2k\pi}{3}\right) \\
                \sin\left(\frac{2k\pi}{3}\right) \\
                c
            \end{pmatrix}
        \]
        for some positive constant $c$.

        For $i\neq j$, the angle between $\mathbf{v}_i$ and $\mathbf{v}_j$ is $120^\circ$, so
        \[
            \mathbf{v}_i\cdot\mathbf{v}_j
            =
            \cos\left(\frac{2\pi}{3}\right)
            =
            -\frac{1}{2}.
        \]
        Therefore,
        \[
            \mathbf{w}_i\cdot\mathbf{w}_j
            =
            \mathbf{v}_i\cdot\mathbf{v}_j+c^2
            =
            -\frac{1}{2}+c^2.
        \]
        To make the vectors mutually orthogonal, we need
        \[
            -\frac{1}{2}+c^2=0,
        \]
        so
        \[
            c=\frac{1}{\sqrt{2}}.
        \]

        Hence one possible choice is
        \[
            \boxed{
            \mathbf{w}_k
            =
            \begin{pmatrix}
                \cos\left(\frac{2k\pi}{3}\right) \\
                \sin\left(\frac{2k\pi}{3}\right) \\
                \frac{1}{\sqrt{2}}
            \end{pmatrix}
            \quad
            (k=0,1,2)
            }.
        \]
        These vectors all have squared length
        \[
            1+\frac{1}{2}
            =
            \frac{3}{2},
        \]
        have positive $z$-coordinates, project to the given vectors $\mathbf{v}_k$, and are mutually orthogonal.
    \end{solution}

    \begin{problem}{2015.4}
        This French mathematician was born in 1973 and received the Fields Medal in 2010 at the age of $36$. On his website, he described himself as ``Math\'ematicien, Professeur de l'Universit\'e de Lyon, Directeur de l'Institut Henri Poincar\'e.'' He is also famous for his stylish clothing.

        Together with Cl\'ement Mouhot, he completed a theorem that played a central role in his Fields Medal-winning work. He later wrote a book about the process of developing this theorem, and the title of the book is also one of his nicknames.

        Write the name of this mathematician, using the correct French spelling, and the title of the book.
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is C\'edric Villani.

        Villani received the Fields Medal in 2010 for his work on nonlinear Landau damping and the Boltzmann equation. He is also widely recognized for his distinctive style, often including a cravat and a spider brooch.

        The book mentioned in the problem is \textit{Th\'eor\`eme vivant}, which may be translated as \textit{Living Theorem}. It describes the process through which Villani and Cl\'ement Mouhot developed the theorem that contributed to Villani's Fields Medal-winning work.

        Therefore, the answer is $\boxed{\text{C\'edric Villani, \textit{Th\'eor\`eme vivant}}}$.
    \end{solution}

    \begin{problem}{2015.5}
        On the front side of the Fields Medal, there is a profile of Archimedes together with the phrase by Marcus Manilius:
        \[
            \text{``TRANSIRE SUUM PECTUS MUNDOQUE POTIRI.''}
        \]
        On the back side, there is another inscription together with the two solid figures that were engraved on Archimedes' tomb.

        Identify the two solid figures and determine their volume ratio.
    \end{problem}

    \begin{solution}
        The two solid figures are a sphere and its circumscribed cylinder.

        Suppose the sphere has radius $R$. Then the volume of the sphere is
        \[
            V_S=\frac{4}{3}\pi R^3.
        \]
        The circumscribed cylinder has radius $R$ and height $2R$, so its volume is
        \[
            V_C=\pi R^2\cdot 2R=2\pi R^3.
        \]
        Therefore,
        \[
            V_S:V_C
            =
            \frac{4}{3}\pi R^3:2\pi R^3
            =
            2:3.
        \]

        Thus the two figures are a sphere and its circumscribed cylinder, and the volume ratio is $\boxed{2:3}$.
    \end{solution}

    \begin{problem}{2015.6}
        Let $f:\R\to\R$ be a periodic function with period $2\pi$. For $x\in(0,2\pi]$, define
        \[
            f(x)
            =
            \left\lceil
                \max(3,|\pi-x|)-3
            \right\rceil,
        \]
        and extend $f$ periodically, so that
        \[
            f(x+2\pi)=f(x)
        \]
        for all $x\in\R$. Determine whether the following limit converges:
        \[
            \lim_{n\to\infty} f(n).
        \]

        \begin{hint}
            Use $3.14<\pi<3.145$.
        \end{hint}
    \end{problem}

    \begin{solution}
        For $x\in(0,2\pi]$, the function $f$ takes only the values $0$ and $1$. More precisely,
        \[
            f(x)=1
        \]
        when
        \[
            0<x<\pi-3
            \qquad\text{or}\qquad
            \pi+3<x\leq 2\pi,
        \]
        and
        \[
            f(x)=0
        \]
        when
        \[
            \pi-3\leq x\leq \pi+3.
        \]
        Thus \(f\) is a \(2\pi\)-periodic function which is \(1\) near multiples of \(2\pi\), and \(0\) on the large middle interval.

        Since $1/(2\pi)$ is irrational, the fractional parts
        \[
            \left\{\frac{n}{2\pi}\right\},
            \qquad n=1,2,3,\dots,
        \]
        are dense in $[0,1]$. Equivalently, the representatives of the integers
        modulo $2\pi$ are dense in $(0,2\pi]$. Hence there are infinitely many
        positive integers $n$ whose representatives modulo $2\pi$ lie in
        \[
            (0,\pi-3)\cup(\pi+3,2\pi],
        \]
        and for these integers we have
        \[
            f(n)=1.
        \]
        There are also infinitely many positive integers \(n\) whose residues modulo \(2\pi\) lie in
        \[
            (\pi-3,\pi+3),
        \]
        and for these integers we have
        \[
            f(n)=0.
        \]

        Therefore, the sequence \(f(n)\) has a subsequence equal to \(1\) and another subsequence equal to \(0\). Hence \(f(n)\) does not converge as \(n\to\infty\).

        Therefore, the sequence $(f(n))$ does not converge, and the answer is $\boxed{\text{the limit does not exist}}$.
    \end{solution}

    \begin{problem}{2015.7}
        This person is believed to have lived approximately from $624$ BCE to $546$ BCE. He is often regarded as one of the earliest figures worthy of being called a mathematician. He introduced the method of proof to geometric facts that had been known empirically through land surveying, helping to lay the foundations of geometry that were later systematized in Euclid's \textit{Elements}.

        The theorems attributed to him include:
        \begin{enumerate}
            \item[(a)] Vertical angles are equal.
            \item[(b)] The base angles of an isosceles triangle are equal.
            \item[(c)] Triangle congruence criteria such as side-angle-side and angle-side-angle.
            \item[(d)] An angle inscribed in a semicircle is a right angle.
        \end{enumerate}
        Identify this person.
    \end{problem}

    \begin{solution}
        The person described in the problem is Thales of Miletus.

        Thales is traditionally regarded as one of the earliest Greek mathematicians. He is especially important because he is associated not merely with observing geometric facts, but with proving them. Several elementary geometric results are attributed to him, including the theorem that an angle inscribed in a semicircle is a right angle.

        Therefore, the answer is $\boxed{\text{Thales of Miletus}}$.
    \end{solution}

    \begin{problem}{2015.8}
        Determine which of the following mathematicians was born earliest.
        \[
            \begin{array}{ll}
                (a) & \text{Henri Poincar\'e} \\
                (b) & \text{Karl Weierstrass} \\
                (c) & \text{Julius Wilhelm Richard Dedekind} \\
                (d) & \text{Andrei Andreyevich Markov} \\
                (e) & \text{Georg Cantor}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        We compare their years of birth:
        \[
            \begin{array}{ll}
                \text{Karl Weierstrass} & 1815, \\
                \text{Julius Wilhelm Richard Dedekind} & 1831, \\
                \text{Georg Cantor} & 1845, \\
                \text{Henri Poincar\'e} & 1854, \\
                \text{Andrei Andreyevich Markov} & 1856.
            \end{array}
        \]
        Among these mathematicians, the one born earliest is Karl Weierstrass.

        Therefore, the answer is $\boxed{\text{Karl Weierstrass}}$.
    \end{solution}

    \begin{problem}{2015.9}
        The first International Congress of Mathematicians at which the Fields Medal was awarded was held in 1936. Determine the country in which it was held.
    \end{problem}

    \begin{solution}
        The Fields Medal was first awarded at the 1936 International Congress of Mathematicians. The 1936 ICM was held in Oslo, Norway.

        Therefore, the answer is $\boxed{\text{Norway}}$.
    \end{solution}

    \begin{problem}{2015.10}
        Find the sum of the distinct prime numbers appearing in the prime factorization of $15!$.
    \end{problem}

    \begin{solution}
        The prime numbers appearing in the prime factorization of $15!$ are exactly the primes less than or equal to $15$:
        \[
            2,\quad 3,\quad 5,\quad 7,\quad 11,\quad 13.
        \]
        Therefore, their sum is
        \[
            2+3+5+7+11+13=41.
        \]
        Therefore, the answer is $\boxed{41}$.
    \end{solution}

    \begin{problem}{2015.11}
        In three-dimensional space, suppose that the lengths of the orthogonal projections of a line segment $\ell$ onto the $yz$-plane, the $zx$-plane, and the $xy$-plane are $5$, $5$, and $6$, respectively. Determine the length of $\ell$.
    \end{problem}

    \begin{solution}
        Let the displacement vector of the line segment be
        \[
            \mathbf{v}=(a,b,c).
        \]
        Then the length of $\ell$ is
        \[
            \|\mathbf{v}\|
            =
            \sqrt{a^2+b^2+c^2}.
        \]

        The projection of $\mathbf{v}$ onto the $yz$-plane has length
        \[
            \sqrt{b^2+c^2}=5.
        \]
        Similarly, the projections onto the $zx$-plane and the $xy$-plane give
        \[
            \sqrt{c^2+a^2}=5,
            \qquad
            \sqrt{a^2+b^2}=6.
        \]
        Squaring and adding these three equations, we get
        \[
            (b^2+c^2)+(c^2+a^2)+(a^2+b^2)
            =
            5^2+5^2+6^2.
        \]
        Hence
        \[
            2(a^2+b^2+c^2)=25+25+36=86,
        \]
        so
        \[
            a^2+b^2+c^2=43.
        \]
        Therefore, the length of $\ell$ is
        \[
            \|\mathbf{v}\|=\sqrt{43}.
        \]

        Therefore, the answer is $\boxed{\sqrt{43}}$.
    \end{solution}

    \begin{problem}{2015.12}
        In three-dimensional space, consider the matrix obtained by rotating space by $30^\circ$ around the line passing through the origin and the point $(1,1,2)$. Determine the number of real eigenvalues of this matrix.
    \end{problem}

    \begin{solution}
        A rotation in $\R^3$ around an axis fixes every vector lying on the rotation axis. Therefore, the direction vector
        \[
            \mathbf{v}
            =
            \begin{pmatrix}
                1 \\
                1 \\
                2
            \end{pmatrix}
        \]
        is an eigenvector with eigenvalue $1$.

        On the plane perpendicular to the rotation axis, the matrix acts as a $30^\circ$ rotation. A nontrivial rotation in a two-dimensional plane has no real eigenvalues unless the rotation angle is $0^\circ$ or $180^\circ$. Since the angle here is $30^\circ$, the two remaining eigenvalues are non-real complex conjugates.

        Hence the only real eigenvalue is $1$. Therefore, the number of real eigenvalues is $\boxed{1}$.
    \end{solution}

    \begin{problem}{2015.13}
        A right circular cone has base radius $3$ and height $6$. Determine the maximum possible volume of a cylinder inscribed in this cone.
    \end{problem}

    \begin{solution}
        Let $x$ be the radius of the inscribed cylinder, and let $h$ be its height. By similarity of triangles in the cone, the radius of the horizontal cross-section decreases linearly from $3$ to $0$ as the height increases from $0$ to $6$. Hence
        \[
            x=3-\frac{h}{2}.
        \]
        Equivalently,
        \[
            h=6-2x.
        \]
        Therefore, the volume of the cylinder is
        \[
            V(x)
            =
            \pi x^2h
            =
            \pi x^2(6-2x)
            =
            6\pi x^2-2\pi x^3,
        \]
        where $0\leq x\leq 3$.

        Differentiating, we get
        \[
            V'(x)
            =
            12\pi x-6\pi x^2
            =
            6\pi x(2-x).
        \]
        Thus the critical point in the interval is
        \[
            x=2.
        \]
        At this point,
        \[
            h=6-2\cdot 2=2,
        \]
        so the maximum volume is
        \[
            V(2)
            =
            \pi\cdot 2^2\cdot 2
            =
            8\pi.
        \]

        Therefore, the maximum possible volume is $\boxed{8\pi}$.
    \end{solution}

    \begin{problem}{2015.14}
        Suppose that a polynomial equation of degree $7$ with real coefficients has three non-real complex roots whose absolute values are all different. Determine the number of real roots of the equation.
    \end{problem}

    \begin{solution}
        For a polynomial with real coefficients, every non-real complex root appears together with its complex conjugate.

        Suppose the three given non-real roots are
        \[
            \alpha_1,\quad \alpha_2,\quad \alpha_3,
        \]
        and their absolute values are all different. Then their complex conjugates
        \[
            \overline{\alpha_1},\quad
            \overline{\alpha_2},\quad
            \overline{\alpha_3}
        \]
        are also roots. Since
        \[
            |\alpha_i|=|\overline{\alpha_i}|,
        \]
        and the three given roots have distinct absolute values, these conjugate roots give three additional non-real roots.

        Thus the polynomial has at least $6$ non-real roots. Since the degree is $7$, there is exactly one remaining root, and it must be real.

        Therefore, the equation has $\boxed{1}$ real root.
    \end{solution}

    \begin{problem}{2015.15}
        This mathematician was born in 1882 and died in 1935. When she was being considered for a professorship at the University of Göttingen, there was strong opposition. In response, Hilbert famously argued, ``I do not see that the sex of the candidate is an argument against her admission as a Privatdozent. After all, we are a university, not a bathhouse.''

        She made remarkable contributions to several areas of mathematics, especially modern algebra, and much of twentieth-century abstract algebra was deeply influenced by her work. Identify this mathematician.
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is Emmy Noether.

        Noether made fundamental contributions to modern algebra, especially through her work on rings, ideals, and modules. She is also famous for Noether's theorem, which connects symmetries and conservation laws in physics.

        The biographical details in the problem, including her connection with Göttingen and Hilbert's defense of her academic appointment, point to Emmy Noether.

        Therefore, the answer is $\boxed{\text{Emmy Noether}}$.
    \end{solution}

    \begin{problem}{2015.16}
        Determine the number of distinct terms in the expansion of
        \[
            (x_1+x_2+\cdots+x_7)^4
        \]
        .
    \end{problem}

    \begin{solution}
        Each term in the expansion has the form
        \[
            x_1^{k_1}x_2^{k_2}\cdots x_7^{k_7},
        \]
        where
        \[
            k_1+k_2+\cdots+k_7=4
        \]
        and each \(k_i\) is a nonnegative integer.

        Therefore, the number of distinct terms is the number of nonnegative integer solutions to
        \[
            k_1+k_2+\cdots+k_7=4.
        \]
        By the stars and bars formula, this number is
        \[
            \binom{4+7-1}{7-1}
            =
            \binom{10}{6}
            =
            210.
        \]

        Therefore, the answer is $\boxed{210}$.
    \end{solution}

    \begin{problem}{2015.17}
        Let
        \[
            \mathbf{v}=\mathbf{w}=(1,2,3,4)
        \]
        be row vectors. Consider the $4\times 4$ matrix obtained as the product
        \[
            \mathbf{v}^{\mathsf T}\mathbf{w}.
        \]
        Find the sum and the product of the eigenvalues of this matrix.
    \end{problem}

    \begin{solution}
        The matrix is
        \[
            \mathbf{v}^{\mathsf T}\mathbf{w}
            =
            \begin{pmatrix}
                1 \\
                2 \\
                3 \\
                4
            \end{pmatrix}
            \begin{pmatrix}
                1 & 2 & 3 & 4
            \end{pmatrix}.
        \]
        The sum of the eigenvalues of a square matrix, counted with algebraic multiplicity, is equal to its trace. Therefore,
        \[
            \sum \lambda_i
            =
            \tr(\mathbf{v}^{\mathsf T}\mathbf{w})
            =
            1^2+2^2+3^2+4^2
            =
            30.
        \]

        The product of the eigenvalues is equal to the determinant. Since $\mathbf{v}^{\mathsf T}\mathbf{w}$ is an outer product of two vectors, we have $\rk(\mathbf{v}^{\mathsf T}\mathbf{w})=1$. In particular, as a $4\times 4$ matrix, it is not invertible, so
        \[
            \det(\mathbf{v}^{\mathsf T}\mathbf{w})=0.
        \]
        Hence the product of the eigenvalues is
        \[
            \prod \lambda_i=0.
        \]

        Therefore, the sum and product of the eigenvalues are $\boxed{30\text{ and }0}$, respectively.
    \end{solution}

    \begin{problem}{2015.18}
        Evaluate the infinite series
        \[
            \sum_{n=1}^{\infty}\frac{n}{(n+1)!}.
        \]

        \begin{hint}
            Use
            \[
                \frac{e^x-1}{x}
                =
                \sum_{n=0}^{\infty}\frac{x^n}{(n+1)!},
            \]
            then differentiate and substitute $x=1$.
        \end{hint}
    \end{problem}

    \begin{solution}
        We use a simple telescoping identity:
        \[
            \frac{n}{(n+1)!}
            =
            \frac{1}{n!}-\frac{1}{(n+1)!}.
        \]
        Therefore,
        \[
            \sum_{n=1}^{N}\frac{n}{(n+1)!}
            =
            \sum_{n=1}^{N}
            \left(
                \frac{1}{n!}-\frac{1}{(n+1)!}
            \right)
            =
            1-\frac{1}{(N+1)!}.
        \]
        Taking \(N\to\infty\), we obtain
        \[
            \sum_{n=1}^{\infty}\frac{n}{(n+1)!}
            =
            1.
        \]
        Therefore, the answer is $\boxed{1}$.
    \end{solution}

    \begin{problem}{2017.1}
        Determine all possible positive integers that can occur as the number of faces of a regular polyhedron, that is, a Platonic solid.
    \end{problem}

    \begin{solution}
        There are exactly five Platonic solids:
        \[
            \begin{array}{c|c}
                \text{Solid} & \text{Number of faces} \\
                \hline
                \text{Tetrahedron} & 4 \\
                \text{Cube} & 6 \\
                \text{Octahedron} & 8 \\
                \text{Dodecahedron} & 12 \\
                \text{Icosahedron} & 20
            \end{array}
        \]
        Therefore, the possible numbers of faces of a regular polyhedron are
        \[
            \boxed{4,\ 6,\ 8,\ 12,\ 20}.
        \]
    \end{solution}

    \begin{problem}{2017.2}
        Let $F_n$ be the $n$-th Fibonacci number. Evaluate
        \[
            \lim_{n\to\infty}\frac{F_{n+1}}{F_n}.
        \]
    \end{problem}

    \begin{solution}
        The Fibonacci numbers satisfy
        \[
            F_{n+1}=F_n+F_{n-1}.
        \]
        Suppose
        \[
            L=\lim_{n\to\infty}\frac{F_{n+1}}{F_n}.
        \]
        Then
        \[
            \frac{F_{n+1}}{F_n}
            =
            1+\frac{F_{n-1}}{F_n}.
        \]
        Taking limits gives
        \[
            L=1+\frac{1}{L}.
        \]
        Hence
        \[
            L^2=L+1,
        \]
        so
        \[
            L=\frac{1\pm\sqrt{5}}{2}.
        \]
        Since the Fibonacci numbers are positive for large \(n\), the limit must be positive. Therefore,
        \[
            L=\frac{1+\sqrt{5}}{2}.
        \]

        Therefore, the answer is $\boxed{\dfrac{1+\sqrt{5}}{2}}$.
    \end{solution}

    \begin{problem}{2017.3}
        For a positive integer $n$, define the $n\times n$ matrix $\mathbf{C}$ by
        \[
            c_{ij}
            =
            \binom{i+j-2}{i-1}
            \qquad
            (1\leq i,j\leq n).
        \]
        Now define another $n\times n$ matrix $\mathbf{B}$ by
        \[
            b_{ij}
            =
            \begin{cases}
                c_{ij}, & (i,j)\neq (n,n), \\
                c_{ij}+2016, & (i,j)=(n,n).
            \end{cases}
        \]
        Determine $\det(\mathbf{B})$.
    \end{problem}

    \begin{solution}
        We first compute the determinant of $\mathbf{C}$. By Vandermonde's identity,
        \[
            \binom{i+j-2}{i-1}
            =
            \sum_{k=1}^{\min(i,j)}
            \binom{i-1}{k-1}
            \binom{j-1}{k-1}.
        \]
        Hence
        \[
            \mathbf{C}=\mathbf{L}\mathbf{L}^{\mathsf T},
        \]
        where
        \[
            l_{ik}
            =
            \begin{cases}
                \binom{i-1}{k-1}, & k\leq i, \\
                0, & k>i.
            \end{cases}
        \]
        The matrix $\mathbf{L}$ is lower triangular and all of its diagonal entries are $1$. Therefore,
        \[
            \det(\mathbf{L})=1,
        \]
        and so
        \[
            \det(\mathbf{C})
            =
            \det(\mathbf{L}\mathbf{L}^{\mathsf T})
            =
            (\det(\mathbf{L}))^2
            =
            1.
        \]

        The matrix $\mathbf{B}$ differs from $\mathbf{C}$ only in the $(n,n)$-entry. Since the determinant is linear in each entry, we have
        \[
            \det(\mathbf{B})
            =
            \det(\mathbf{C})
            +
            2016\cdot M_{nn},
        \]
        where $M_{nn}$ is the $(n,n)$-minor of $\mathbf{C}$.

        This minor is the determinant of the upper-left $(n-1)\times(n-1)$ submatrix of $\mathbf{C}$, which has the same form as $\mathbf{C}$ with size $n-1$. Therefore,
        \[
            M_{nn}=1.
        \]
        Hence
        \[
            \det(\mathbf{B})
            =
            1+2016\cdot 1
            =
            2017.
        \]

        Therefore, the answer is $\boxed{2017}$.
    \end{solution}

    \begin{problem}{2017.4}
        In 2017, this prize was awarded to the French mathematician Yves Meyer. The Norwegian Academy of Science and Letters announced on March 21 that Meyer would receive the prize, worth six million Norwegian kroner, in recognition of his decisive contributions to the development of wavelet theory, which is used in areas ranging from gravitational wave detection to movie file compression.

        Identify this prize, which is often called the Nobel Prize of mathematics.
    \end{problem}

    \begin{solution}
        The prize described in the problem is the Abel Prize.

        The Abel Prize is awarded by the Norwegian Academy of Science and Letters and is often described as one of the most prestigious prizes in mathematics. In 2017, it was awarded to Yves Meyer for his fundamental contributions to the mathematical theory of wavelets.

        Therefore, the answer is $\boxed{\text{Abel Prize}}$.
    \end{solution}

    \begin{problem}{2017.5}
        Express the integral
        \[
            \int_0^1 x^m(1-x)^n\dd x
        \]
        in terms of a binomial coefficient closely related to the positive integers $m$ and $n$.
    \end{problem}

    \begin{solution}
        This integral is a beta integral. For positive integers $m$ and $n$, we have
        \[
            \int_0^1 x^m(1-x)^n\dd x
            =
            \frac{m!n!}{(m+n+1)!}.
        \]
        This is the factorial form of the answer.

        To express it using a binomial coefficient, recall that
        \[
            \binom{m+n}{m}
            =
            \frac{(m+n)!}{m!n!}.
        \]
        Hence
        \[
            \frac{m!n!}{(m+n+1)!}
            =
            \frac{1}{m+n+1}\cdot\frac{m!n!}{(m+n)!}
            =
            \frac{1}{(m+n+1)\binom{m+n}{m}}.
        \]
        Therefore,
        \[
            \int_0^1 x^m(1-x)^n\dd x
            =
            \frac{m!n!}{(m+n+1)!}
            =
            \frac{1}{(m+n+1)\binom{m+n}{m}}.
        \]

        Therefore, the answer is $\boxed{\dfrac{m!n!}{(m+n+1)!}
        =
        \dfrac{1}{(m+n+1)\binom{m+n}{m}}}$.
    \end{solution}

    \begin{problem}{2018.1}
        The following people are this year's Fields Medalists. Excluding choice (5), identify the second-oldest medalist among them.
        \[
            \begin{array}{ll}
                (1) & \text{Caucher Birkar} \\
                (2) & \text{Alessio Figalli} \\
                (3) & \text{Peter Scholze} \\
                (4) & \text{Akshay Venkatesh} \\
                (5) & \text{The person who stole Birkar's medal}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The 2018 Fields Medalists were
        \[
            \begin{array}{lll}
                (1) & \text{Caucher Birkar} & \text{born in 1978}, \\
                (2) & \text{Alessio Figalli} & \text{born in 1984}, \\
                (3) & \text{Peter Scholze} & \text{born in 1987}, \\
                (4) & \text{Akshay Venkatesh} & \text{born in 1981}.
            \end{array}
        \]
        Among these four, the second oldest is Akshay Venkatesh. Therefore, the answer is
        \[
            \boxed{(4)}.
        \]

        Choice (5) is a joke referring to the incident in which Caucher Birkar's Fields Medal was stolen shortly after the 2018 Fields Medal ceremony. Birkar later received a replacement medal.
    \end{solution}

    \begin{problem}{2018.2}
        According to William Thurston's geometrization conjecture, a $3$-dimensional manifold can be cut along spheres and tori so that each resulting piece has one of $n$ possible geometric structures. Determine $n$.
    \end{problem}

    \begin{solution}
        The $3$-dimensional geometric structures appearing in Thurston's geometrization conjecture are the eight Thurston geometries:
        \[
            S^3,\quad
            \mathrm{E}^3,\quad
            \mathbb{H}^3,\quad
            S^2 \times \R,\quad
            \mathbb{H}^2 \times \R,\quad
            \widetilde{\SL}_2(\R),\quad
            \mathrm{Nil},\quad
            \mathrm{Sol}.
        \]
        Hence $n=8$, so the answer is
        \[
            \boxed{8}.
        \]
    \end{solution}

    \begin{problem}{2018.3}
        Find the sum of the following infinite series:
        \[
            \frac{1}{2}-\frac{1}{4}+\frac{1}{6}-\frac{1}{8}
            +\frac{1}{10}-\frac{1}{12}+\cdots.
        \]
    \end{problem}

    \begin{solution}
        The given series is one half of the alternating harmonic series:
        \[
            \frac{1}{2}-\frac{1}{4}+\frac{1}{6}-\frac{1}{8}+\cdots
            =
            \frac{1}{2}
            \left(
                1-\frac{1}{2}+\frac{1}{3}-\frac{1}{4}+\cdots
            \right).
        \]
        Recall the Taylor expansion
        \[
            \ln(1+x)
            =
            x-\frac{x^2}{2}+\frac{x^3}{3}-\frac{x^4}{4}+\cdots.
        \]
        Substituting $x=1$, we obtain
        \[
            \ln 2
            =
            1-\frac{1}{2}+\frac{1}{3}-\frac{1}{4}+\cdots.
        \]
        Therefore,
        \[
            \frac{1}{2}-\frac{1}{4}+\frac{1}{6}-\frac{1}{8}+\cdots
            =
            \frac{\ln 2}{2}.
        \]
        Therefore, the answer is $\boxed{\dfrac{\ln 2}{2}}$.
    \end{solution}

    \begin{problem}{2018.4}
        Determine which of the following five Fields Medalists does not have a Korean doctoral student.
        \[
            \begin{array}{ll}
                (1) & \text{William Thurston} \\
                (2) & \text{Maxim Kontsevich} \\
                (3) & \text{Grigory Margulis} \\
                (4) & \text{Terence Tao} \\
                (5) & \text{Efim Zelmanov}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        According to the original answer key, the exceptional mathematician among the five listed Fields Medalists is
        \[
            \boxed{(2)\ \text{Maxim Kontsevich}}.
        \]
    \end{solution}

    \begin{problem}{2019.1}
        Consider the $2019\times 2019$ matrix whose diagonal entries are all $0$ and whose off-diagonal entries are all $1$. Find all eigenvalues of this matrix.
    \end{problem}

    \begin{solution}
        Let $\mathbf{A}$ be the given matrix, and let $\mathbf{J}$ be the $2019\times 2019$ matrix whose entries are all $1$. Then
        \[
            \mathbf{A}=\mathbf{J}-\mathbf{I}.
        \]

        First, consider the vector
        \[
            \mathbf{1}=(1,1,\dots,1)^{\mathsf T}.
        \]
        Since
        \[
            \mathbf{J}\mathbf{1}=2019\mathbf{1},
        \]
        we get
        \[
            \mathbf{A}\mathbf{1}
            =
            (\mathbf{J}-\mathbf{I})\mathbf{1}
            =
            2019\mathbf{1}-\mathbf{1}
            =
            2018\mathbf{1}.
        \]
        Hence $2018$ is an eigenvalue.

        Now take any vector $\mathbf{v}=(v_1,\dots,v_{2019})^{\mathsf T}$ satisfying
        \[
            v_1+\cdots+v_{2019}=0.
        \]
        Then
        \[
            \mathbf{J}\mathbf{v}=\mathbf{0},
        \]
        and therefore
        \[
            \mathbf{A}\mathbf{v}
            =
            (\mathbf{J}-\mathbf{I})\mathbf{v}
            =
            -\mathbf{v}.
        \]
        Thus $-1$ is an eigenvalue. The subspace of vectors whose coordinates sum to $0$ has dimension $2018$, so the multiplicity of $-1$ is $2018$.

        Therefore, the eigenvalues are
        \[
            \boxed{2018 \text{ with multiplicity } 1,\qquad -1 \text{ with multiplicity } 2018}.
        \]
    \end{solution}

    \begin{problem}{2019.2}
        Evaluate the integral
        \[
            \int_{-\infty}^{\infty}
            \frac{8}{1+4x^2}\dd x.
        \]
    \end{problem}

    \begin{solution}
        Let
        \[
            u=2x.
        \]
        Then
        \[
            \dd x=\frac{1}{2}\dd u.
        \]
        Therefore,
        \[
            \int_{-\infty}^{\infty}
            \frac{8}{1+4x^2}\dd x
            =
            \int_{-\infty}^{\infty}
            \frac{8}{1+u^2}\cdot\frac{1}{2}\dd u
            =
            4\int_{-\infty}^{\infty}
            \frac{1}{1+u^2}\dd u.
        \]
        Since
        \[
            \int_{-\infty}^{\infty}
            \frac{1}{1+u^2}\dd u
            =
            \pi,
        \]
        we obtain
        \[
            \int_{-\infty}^{\infty}
            \frac{8}{1+4x^2}\dd x
            =
            4\pi.
        \]

        Therefore, the answer is $\boxed{4\pi}$.
    \end{solution}

    \begin{problem}{2019.3}
        Compute the value of $\pi+e$ up to five decimal places.
    \end{problem}

    \begin{solution}
        Using the approximations
        \[
            \pi\approx 3.14159265,
            \qquad
            e\approx 2.71828183,
        \]
        we get
        \[
            \pi+e
            \approx
            3.14159265+2.71828183
            =
            5.85987448.
        \]
        Therefore, up to five decimal places,
        \[
            \pi+e\approx 5.85987.
        \]

        Therefore, the answer is $\boxed{5.85987}$.
    \end{solution}

    \begin{problem}{2020.1}
        The International Congress of Mathematicians, abbreviated as ICM, is a congress for mathematicians from around the world. It is organized every four years by the International Mathematical Union and may be thought of as a kind of Olympic Games for mathematicians. The 2014 ICM was held in South Korea, and the 2018 ICM was held in Brazil. As of 2019, determine the country in which the 2022 ICM was scheduled to be held.
    \end{problem}

    \begin{solution}
        ICM stands for the International Congress of Mathematicians. It is the main international congress in mathematics and is organized every four years by the International Mathematical Union.

        The 2014 ICM was held in Seoul, South Korea, and the 2018 ICM was held in Rio de Janeiro, Brazil. At the reference date specified in the problem, the 2022 ICM was scheduled to be held in Saint Petersburg, Russia. Therefore, the answer is $\boxed{\text{Russia}}$.

        The congress was later changed to a virtual event, with the 2022 awards ceremony held in Helsinki, Finland. The answer above therefore records the planned host country at the date specified in the original problem.
    \end{solution}

    \begin{problem}{2020.2}
        A fair six-sided die has the numbers $-2,0,2,3,4,5$ written on its faces. The die is rolled $10$ times, and $X$ is defined to be the product of the $10$ numbers obtained. Determine $\Exp[X]$.
    \end{problem}

    \begin{solution}
        Let $Y$ be the number obtained from one roll of the die. Since all six faces are equally likely,
        \[
            \Exp[Y]
            =
            \frac{-2+0+2+3+4+5}{6}
            =
            2.
        \]
        Now let $Y_1,Y_2,\dots,Y_{10}$ be the numbers obtained from the $10$ rolls. Then
        \[
            X=Y_1Y_2\cdots Y_{10}.
        \]
        Since the rolls are independent, the random variables $Y_1,Y_2,\dots,Y_{10}$ are independent. Therefore,
        \[
            \Exp[X]
            =
            \Exp[Y_1Y_2\cdots Y_{10}]
            =
            \Exp[Y_1]\Exp[Y_2]\cdots\Exp[Y_{10}].
        \]
        All the $Y_i$ have the same distribution as $Y$, so
        \[
            \Exp[X]
            =
            \left(\Exp[Y]\right)^{10}
            =
            2^{10}
            =
            1024.
        \]
        Therefore, the answer is $\boxed{1024}$.
    \end{solution}

    \begin{problem}{2020.3}
        Determine which of the following Fields Medalists was not serving as a professor at the time.
        \[
            \begin{array}{ll}
                (1) & \text{Cédric Villani} \\
                (2) & \text{Elon Lindenstrauss} \\
                (3) & \text{Peter Scholze} \\
                (4) & \text{Martin Hairer} \\
                (5) & \text{Terence Tao}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        According to the original answer key and the temporal context of the problem, Cédric Villani was then serving in French politics rather than in a professorial position. The other choices were serving as professors.

        Therefore, the answer is $\boxed{(1)\ \text{Cédric Villani}}$.
    \end{solution}

    \begin{problem}{2020.4}
        In a three-dimensional gravitational field, when an object moves from a point $\mathbf{a}$ to another point $\mathbf{b}$, the change in its potential energy is independent of the path taken from $\mathbf{a}$ to $\mathbf{b}$. Determine which of the following mathematical facts is least directly related to this physical phenomenon.
        \[
            \begin{array}{ll}
                (1) & \text{A curl-free vector field is the gradient of some potential function.} \\
                (2) & \text{Stokes' theorem} \\
                (3) & \text{The Laplacian of } \dfrac{1}{|\mathbf{x}|} \text{ is a delta function, up to a constant factor.} \\
                (4) & \text{The curl of a gradient is identically zero.}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The statement that the change in potential energy is independent of the path means that the gravitational field is a conservative vector field.

        Let $\mathbf{F}$ be the force field. The change in potential energy from $\mathbf{a}$ to $\mathbf{b}$ is given by
        \[
            \Delta U
            =
            -\int_{\mathbf{a}}^{\mathbf{b}} \mathbf{F}\cdot\dd\mathbf{r}.
        \]
        For this quantity to be independent of the path, the force field should be expressible as the negative gradient of a potential function:
        \[
            \mathbf{F}=-\nabla U.
        \]
        Thus choice (1) is directly related to the phenomenon.

        Also, the curl of a gradient is always zero:
        \[
            \nabla \times \nabla U=\mathbf{0}.
        \]
        Hence choice (4) is also directly related. Stokes' theorem says that
        \[
            \oint_{\partial S}\mathbf{F}\cdot\dd\mathbf{r}
            =
            \iint_S (\nabla\times \mathbf{F})\cdot \mathbf{n}\dd S.
        \]
        Therefore, if $\nabla\times\mathbf{F}=\mathbf{0}$, then the line integral around a closed curve is zero, which is another way to understand path independence. Thus choice (2) is also related.

        On the other hand,
        \[
            \Delta\left(\frac{1}{|\mathbf{x}|}\right)
            =
            -4\pi\delta(\mathbf{x})
        \]
        is related to the fundamental solution of the Laplacian and to Poisson's equation for gravitational potential. Although it is related to gravity, it is not the main mathematical reason why the potential energy change is independent of path. Therefore, the answer is $\boxed{(3)}$.
    \end{solution}

    \begin{problem}{2020.5}
        Determine which of the following mathematics-related academic institutions is a degree-granting school.
        \[
            \begin{array}{ll}
                (1) & \text{Institut des Hautes Études Scientifiques, IHES} \\
                (2) & \text{École normale supérieure} \\
                (3) & \text{Mathematical Sciences Research Institute, MSRI} \\
                (4) & \text{Institute for Advanced Study, IAS} \\
                (5) & \text{Collège de France}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        École normale supérieure is one of France's most prestigious higher education institutions. It operates educational programs and is naturally regarded as a degree-granting school.

        By contrast, IHES, MSRI, and IAS are research institutes rather than degree-granting schools. Collège de France is also a famous academic institution that offers public lectures, but it does not grant degrees.

        Therefore, the answer is $\boxed{(2)\ \text{École normale supérieure}}$.
    \end{solution}

    \begin{problem}{2021.1}
        Determine which of $A$ and $B$ is larger, where
        \[
            A
            =
            1+\frac{1}{1+\frac{1}{1-\frac{1}{1+4}}},
            \qquad
            B
            =
            1+\frac{1}{1+\frac{1}{1+\frac{1}{1+3}}}.
        \]
    \end{problem}

    \begin{solution}
        First, we compute $A$:
        \[
            A
            =
            1+\frac{1}{1+\frac{1}{1-\frac{1}{5}}}
            =
            1+\frac{1}{1+\frac{1}{\frac{4}{5}}}
            =
            1+\frac{1}{1+\frac{5}{4}}
            =
            1+\frac{1}{\frac{9}{4}}
            =
            1+\frac{4}{9}
            =
            \frac{13}{9}.
        \]
        Next, we compute $B$:
        \[
            B
            =
            1+\frac{1}{1+\frac{1}{1+\frac{1}{4}}}
            =
            1+\frac{1}{1+\frac{1}{\frac{5}{4}}}
            =
            1+\frac{1}{1+\frac{4}{5}}
            =
            1+\frac{1}{\frac{9}{5}}
            =
            1+\frac{5}{9}
            =
            \frac{14}{9}.
        \]
        Since
        \[
            \frac{13}{9}<\frac{14}{9},
        \]
        we have $A<B$. Therefore, the answer is $\boxed{B}$.
    \end{solution}

    \begin{problem}{2021.2}
        Let
        \[
            A(n)
            =
            \frac{1}{\sqrt{5}}
            \left(
                \frac{1+\sqrt{5}}{2}
            \right)^{n+1}.
        \]
        Determine the integer closest to $A(7)$.
    \end{problem}

    \begin{solution}
        Recall Binet's formula for the Fibonacci sequence:
        \[
            F_n
            =
            \frac{\varphi^n-\psi^n}{\sqrt{5}},
            \qquad
            \varphi=\frac{1+\sqrt{5}}{2},
            \qquad
            \psi=\frac{1-\sqrt{5}}{2}.
        \]
        Here
        \[
            A(7)
            =
            \frac{\varphi^8}{\sqrt{5}}.
        \]
        By Binet's formula,
        \[
            F_8
            =
            \frac{\varphi^8-\psi^8}{\sqrt{5}}.
        \]
        Since $|\psi|<1$, the term $\dfrac{\psi^8}{\sqrt{5}}$ is very small. Thus
        \[
            A(7)
            =
            \frac{\varphi^8}{\sqrt{5}}
            =
            F_8+\frac{\psi^8}{\sqrt{5}}
        \]
        is very close to $F_8$.

        The Fibonacci numbers are
        \[
            F_0=0,\quad F_1=1,\quad F_2=1,\quad F_3=2,\quad F_4=3,\quad
            F_5=5,\quad F_6=8,\quad F_7=13,\quad F_8=21.
        \]
        Therefore, the integer closest to $A(7)$ is $\boxed{21}$.
    \end{solution}

    \begin{problem}{2021.3}
        Choose all groups among the following that are not finitely generated:
        \[
            \begin{array}{ll}
                (1) & \text{The fundamental group of an open unit disk with finitely many punctures.} \\
                (2) & \SL_2(\Z), \text{ the group of } 2\times 2 \text{ integer matrices with determinant } 1. \\
                (3) & \C^{\times}, \text{ the group of nonzero complex numbers under multiplication.} \\
                (4) & \text{The Monster group, the largest sporadic finite simple group.}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        First, consider an open unit disk with finitely many punctures. If the removed points are $p_1,\dots,p_n$, then the fundamental group is isomorphic to the free group on $n$ generators:
        \[
            \pi_1\left(D\setminus\{p_1,\dots,p_n\}\right)\cong F_n.
        \]
        Hence the group in choice (1) is finitely generated.

        The group $\SL_2(\Z)$ is also finitely generated. For example, it is generated by the two matrices
        \[
            \mathbf{S}=
            \begin{pmatrix}
                0 & -1 \\
                1 & 0
            \end{pmatrix},
            \qquad
            \mathbf{T}=
            \begin{pmatrix}
                1 & 1 \\
                0 & 1
            \end{pmatrix}.
        \]
        Thus choice (2) is finitely generated.

        On the other hand, $\C^{\times}$ is an uncountable abelian group. Every finitely generated group is countable, so $\C^{\times}$ cannot be finitely generated. Therefore, choice (3) is not finitely generated.

        Finally, the Monster group is a finite group. Every finite group is finitely generated, since one may simply take all of its elements as generators. Hence choice (4) is finitely generated.

        Therefore, the only group in the list that is not finitely generated is $\boxed{(3)\ \C^{\times}}$.
    \end{solution}

    \begin{problem}{2021.4}
        Let
        \[
            \mathbf{A}
            =
            \begin{pmatrix}
                1 & 2 & 3 \\
                1 & 2 & 3 \\
                1 & 2 & 3
            \end{pmatrix}.
        \]
        Find all eigenvalues of $\mathbf{A}$ together with their multiplicities. Also determine whether $\mathbf{A}$ is diagonalizable.
    \end{problem}

    \begin{solution}
        We can write $\mathbf{A}$ as
        \[
            \mathbf{A}
            =
            \begin{pmatrix}
                1 \\
                1 \\
                1
            \end{pmatrix}
            \begin{pmatrix}
                1 & 2 & 3
            \end{pmatrix}.
        \]
        Thus $\rk(\mathbf{A})=1$.

        First, let
        \[
            \mathbf{v}
            =
            \begin{pmatrix}
                1 \\
                1 \\
                1
            \end{pmatrix}.
        \]
        Then
        \[
            \mathbf{A}\mathbf{v}
            =
            \begin{pmatrix}
                6 \\
                6 \\
                6
            \end{pmatrix}
            =
            6\mathbf{v}.
        \]
        Hence $6$ is an eigenvalue of $\mathbf{A}$.

        Since $\rk(\mathbf{A})=1$, its nullity is $2$. Therefore, the eigenspace corresponding to the eigenvalue $0$ has dimension $2$. Indeed,
        \[
            \mathbf{A}
            \begin{pmatrix}
                x \\
                y \\
                z
            \end{pmatrix}
            =
            \begin{pmatrix}
                x+2y+3z \\
                x+2y+3z \\
                x+2y+3z
            \end{pmatrix},
        \]
        so
        \[
            E_0
            =
            \left\{
                \begin{pmatrix}
                    x \\
                    y \\
                    z
                \end{pmatrix}
                \in \R^3
                :
                x+2y+3z=0
            \right\}.
        \]
        Thus the geometric multiplicity of $0$ is $2$.

        Now
        \[
            \tr(\mathbf{A})=1+2+3=6.
        \]
        Since the sum of the eigenvalues, counted with algebraic multiplicity, is equal to $\tr(\mathbf{A})$, the eigenvalues are
        \[
            6,\quad 0,\quad 0.
        \]
        Hence the algebraic multiplicities are
        \[
            6:1,
            \qquad
            0:2.
        \]

        Finally, the eigenspace for $6$ has dimension $1$, and the eigenspace for $0$ has dimension $2$. The total dimension of the eigenspaces is therefore
        \[
            1+2=3.
        \]
        Hence there exists a basis of $\R^3$ consisting of eigenvectors of $\mathbf{A}$. Therefore, $\mathbf{A}$ is diagonalizable.

        Thus the eigenvalues are $\boxed{6,0,0}$, with algebraic multiplicities $1$ and $2$ respectively, and $\boxed{\mathbf{A}\text{ is diagonalizable}}$.
    \end{solution}

    \begin{problem}{2022.1}
        Determine the largest of the following numbers.
        \[
            \begin{array}{ll}
                (1) & e^\pi \\
                (2) & \pi^e \\
                (3) & 7\pi \\
                (4) & \binom{7}{2} \\
                (5) & 4!
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        We compare the values directly:
        \[
            e^\pi \approx 23.1407,
            \qquad
            \pi^e \approx 22.4592,
            \qquad
            7\pi \approx 21.9911.
        \]
        Also,
        \[
            \binom{7}{2}
            =
            \frac{7\cdot 6}{2}
            =
            21,
            \qquad
            4!
            =
            24.
        \]
        Therefore, the largest number among the choices is $\boxed{(5)\ 4!}$.
    \end{solution}

    \begin{problem}{2022.2}
        Three people store a jointly owned treasure in a safe and want to lock it. They want no single person to be able to open the safe alone, but any two or more people should always be able to open it together. This can be done by putting $n$ locks on the safe, making $m$ copies of the key for each lock, and distributing the keys appropriately among the three people. Find $n$ and $m$ so that the number of locks $n$ is minimized.
    \end{problem}

    \begin{solution}
        Let the three people be $A$, $B$, and $C$.

        Since any two people should be able to open the safe, the key to each lock must be given to at least two people. If a certain lock had its key given to only one person, then the other two people would not be able to open that lock together.

        Thus, for each lock, exactly one of the three people should not receive the key. On the other hand, no single person should be able to open the safe alone. Therefore, each person must be missing the key to at least one lock.

        Hence we need at least three locks: one lock whose key is not given to $A$, one lock whose key is not given to $B$, and one lock whose key is not given to $C$. This lower bound is achievable by distributing the keys as follows:
        \[
            \begin{array}{c|ccc}
                & A & B & C \\
                \hline
                L_A & \times & \checkmark & \checkmark \\
                L_B & \checkmark & \times & \checkmark \\
                L_C & \checkmark & \checkmark & \times
            \end{array}
        \]
        That is, the key to $L_A$ is given to $B$ and $C$, the key to $L_B$ is given to $A$ and $C$, and the key to $L_C$ is given to $A$ and $B$.

        Then each person is unable to open one of the locks, so no one can open the safe alone. However, any two people together can open all three locks. Therefore, the minimum is achieved by $\boxed{n=3,\ m=2}$.
    \end{solution}

    \begin{problem}{2022.3}
        Evaluate the integral
        \[
            \int_{-\infty}^{\infty}
            \frac{3}{(x-2)^2+7}\dd x.
        \]
    \end{problem}

    \begin{solution}
        Let
        \[
            u=x-2.
        \]
        Then the limits of integration remain $-\infty$ and $\infty$, so
        \[
            \int_{-\infty}^{\infty}
            \frac{3}{(x-2)^2+7}\dd x
            =
            3\int_{-\infty}^{\infty}
            \frac{1}{u^2+7}\dd u.
        \]
        We use the standard formula
        \[
            \int_{-\infty}^{\infty}
            \frac{1}{u^2+a^2}\dd u
            =
            \frac{\pi}{a}.
        \]
        Here $a=\sqrt{7}$, and therefore
        \[
            3\int_{-\infty}^{\infty}
            \frac{1}{u^2+7}\dd u
            =
            3\cdot\frac{\pi}{\sqrt{7}}
            =
            \frac{3\pi}{\sqrt{7}}.
        \]
        Therefore, the answer is $\boxed{\dfrac{3\pi}{\sqrt{7}}}$.
    \end{solution}

    \begin{problem}{2022.4}
        A couple has two children.
        \begin{enumerate}
            \item[(a)] Given that at least one child is a daughter, determine the probability that both children are daughters.
            \item[(b)] Given that a specified child is a daughter named Youngmi, determine the probability that the other child is also a daughter.
        \end{enumerate}
    \end{problem}

    \begin{solution}
        Let $G$ denote a girl and $B$ denote a boy. If we distinguish the two children by birth order, then the possible gender combinations are
        \[
            GG,\quad GB,\quad BG,\quad BB.
        \]

        For part (a), the condition that one child is a daughter means that at least one child is a girl. Thus the possible cases are
        \[
            GG,\quad GB,\quad BG.
        \]
        Among these three cases, only $GG$ has the property that the other child is also a daughter. Therefore,
        \[
            \Prob(\text{both are daughters}\mid \text{at least one is a daughter})
            =
            \frac{1}{3}.
        \]

        For part (b), the information that one child is a daughter named Youngmi is more specific. In the intended interpretation of the problem, this information identifies one particular child. Then the gender of the other child is still equally likely to be $G$ or $B$. Hence
        \[
            \Prob(\text{the other child is a daughter})
            =
            \frac{1}{2}.
        \]

        Therefore, the answers are $\boxed{\text{(a) } \dfrac{1}{3},\ \text{(b) } \dfrac{1}{2}}$.
    \end{solution}

    \begin{problem}{2022.5}
        Determine which of the following conjectures was not solved by Professor June Huh.
        \[
            \begin{array}{ll}
                (1) & \text{Read's conjecture} \\
                (2) & \text{Brylawski's conjecture} \\
                (3) & \text{Poincaré conjecture} \\
                (4) & \text{Rota's conjecture} \\
                (5) & \text{Okounkov's conjecture}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Professor June Huh is known for connecting combinatorics with algebraic geometry and using these ideas to solve several major conjectures in combinatorics. His work is related to conjectures such as Read's conjecture, Brylawski's conjecture, Rota's conjecture, and Okounkov's conjecture.

        On the other hand, the Poincaré conjecture is a famous problem in $3$-dimensional topology. It was proved by Grigori Perelman using Ricci flow, not by June Huh.

        Therefore, the conjecture in the list that was not solved by June Huh is
        \[
            \boxed{(3)\ \text{Poincaré conjecture}}.
        \]
    \end{solution}

    \begin{problem}{2022.6}
        This mathematician, often called the father of linear programming, is famous for solving difficult problems while he was a doctoral student. One day, while studying at UC Berkeley, he arrived late to a class. The statistics professor Jerzy Neyman had written two famous unsolved problems on the blackboard, but the student thought they were homework problems. He took them home, solved them, and submitted complete solutions a few days later, thinking only that he had turned in the homework late. Identify this mathematician.
    \end{problem}

    \begin{solution}
        The mathematician described in the problem is George Dantzig.

        George Dantzig is known as one of the founders of linear programming and as the creator of the simplex method. For this reason, he is often called the father of linear programming.

        The anecdote in the problem is also a famous story about Dantzig. While he was a doctoral student at UC Berkeley, he arrived late to Jerzy Neyman's statistics class and copied down two problems written on the blackboard, believing that they were homework problems. He later submitted solutions, not realizing that the problems were actually open problems in statistics.

        Therefore, the answer is $\boxed{\text{George Dantzig}}$.
    \end{solution}

    \begin{problem}{2022.7}
        Two different coins are each tossed three times. Let $X$ and $Y$ be the numbers of heads obtained from the two coins, respectively. Determine $\Prob(X=Y)$.
    \end{problem}

    \begin{solution}
        For each coin, the number of heads obtained in three tosses follows the binomial distribution
        \[
            X,Y\sim \Bin\left(3,\frac{1}{2}\right).
        \]
        Also, $X$ and $Y$ are independent. Therefore,
        \[
            \Prob(X=Y)
            =
            \sum_{k=0}^{3}\Prob(X=k)\Prob(Y=k).
        \]
        For each $k=0,1,2,3$, we have
        \[
            \Prob(X=k)
            =
            \binom{3}{k}\left(\frac{1}{2}\right)^3,
        \]
        and the same formula holds for $Y$. Hence
        \[
            \Prob(X=Y)
            =
            \sum_{k=0}^{3}
            \left(
                \binom{3}{k}\left(\frac{1}{2}\right)^3
            \right)^2
            =
            \frac{1}{64}
            \sum_{k=0}^{3}
            \binom{3}{k}^2.
        \]
        Thus
        \[
            \Prob(X=Y)
            =
            \frac{1}{64}
            \left(
                \binom{3}{0}^2+
                \binom{3}{1}^2+
                \binom{3}{2}^2+
                \binom{3}{3}^2
            \right)
            =
            \frac{1+9+9+1}{64}
            =
            \frac{20}{64}
            =
            \frac{5}{16}.
        \]
        Therefore, the answer is $\boxed{\dfrac{5}{16}}$.
    \end{solution}

    \begin{problem}{2023.1}
        Determine the year in which the Fields Medal, received by Professor June Huh, was first awarded.
    \end{problem}

    \begin{solution}
        The Fields Medal is one of the most prestigious awards in mathematics, and it is awarded by the International Mathematical Union. Professor June Huh received the Fields Medal in 2022.

        The Fields Medal was first awarded in 1936. The first two recipients were Lars Ahlfors and Jesse Douglas.

        Therefore, the answer is $\boxed{1936}$.
    \end{solution}

    \begin{problem}{2023.2}
        Consider the $6\times 6$ square grid with vertices $(0,0)$, $(6,0)$, $(0,6)$, and $(6,6)$. A path starts at $(0,0)$ and ends at $(6,6)$, moving only one unit to the right or one unit upward at each step. Determine the number of such paths that never go above the line $x=y$. Equivalently, determine the number of such paths that stay in the region
        \[
            \{(x,y)\in\R^2:y\leq x\}.
        \]
    \end{problem}

    \begin{solution}
        To go from $(0,0)$ to $(6,6)$, a path must consist of $6$ right steps and $6$ upward steps. Without the condition $y\leq x$, the total number of such paths is
        \[
            \binom{12}{6}.
        \]

        We now count the paths that do go above the line $y=x$. By the reflection principle, the number of paths from $(0,0)$ to $(6,6)$ that at some point satisfy $y>x$ is
        \[
            \binom{12}{5}.
        \]
        Therefore, the number of paths that always stay in the region $y\leq x$ is
        \[
            \binom{12}{6}-\binom{12}{5}.
        \]
        Computing this gives
        \[
            \binom{12}{6}-\binom{12}{5}
            =
            924-792
            =
            132.
        \]
        Therefore, the answer is $\boxed{132}$.
    \end{solution}

    \begin{problem}{2023.3}
        A point on the surface of the Earth has the following property: starting from that point, one moves $5\mathrm{km}$ south, then $5\mathrm{km}$ east, and then $5\mathrm{km}$ north, and returns to the original point. Determine the number of such points on the surface of the Earth.

        Assume that the surface of the Earth is a perfect two-dimensional sphere, and that the North Pole and the South Pole are antipodal points. At the North Pole, the directions north, east, and west are not defined, and at the South Pole, the directions south, east, and west are not defined. Starting points for which one of these undefined directions occurs during the movement are excluded.
    \end{problem}

    \begin{solution}
        The North Pole is one such point. If we start at the North Pole, move $5\mathrm{km}$ south, then move $5\mathrm{km}$ east along a latitude circle, and finally move $5\mathrm{km}$ north, we return to the North Pole.

        There are also many such points near the South Pole. Let $Q$ be the point reached after moving $5\mathrm{km}$ south from the starting point. Suppose that the latitude circle passing through $Q$ has circumference
        \[
            \frac{5}{n}\mathrm{km}
        \]
        for some positive integer $n$. Then moving $5\mathrm{km}$ east from $Q$ goes exactly $n$ full times around that latitude circle and returns to $Q$. After that, moving $5\mathrm{km}$ north returns to the original starting point.

        Near the South Pole, the circumferences of latitude circles approach $0$. Hence, for every positive integer $n$, there is a latitude circle near the South Pole whose circumference is exactly $\dfrac{5}{n}\mathrm{km}$. For each such latitude circle, every point lying $5\mathrm{km}$ north of it gives a valid starting point.

        Therefore, there are not just one but infinitely many such points on the Earth's surface. In fact, this construction gives entire latitude circles of valid starting points, so there are uncountably many such points. For the usual quiz answer, the essential conclusion is
        \[
            \boxed{\text{infinitely many}}.
        \]
    \end{solution}

    \begin{problem}{2023.4}
        Let $f:(-1,1)\to\R$ be a continuous function. For each $\lambda>0$, define
        \[
            f_\lambda(x)
            :=
            \lambda f\left(\frac{x}{\lambda}\right),
        \]
        where $f_\lambda$ is defined for $x\in(-\lambda,\lambda)$.

        Suppose that for every sequence $\lambda_i\to\infty$, the functions $f_{\lambda_i}$ converge locally uniformly to $0$ on $\R$. Determine whether the following statement is true or false:
        \[
            f \text{ is differentiable at } 0 \text{ and } f'(0)=0.
        \]
    \end{problem}

    \begin{solution}
        We prove that the statement is true.

        First, evaluate at $x=0$. For every sequence $\lambda_i\to\infty$, we have
        \[
            f_{\lambda_i}(0)
            =
            \lambda_i f(0).
        \]
        Since $f_{\lambda_i}$ converges locally uniformly to $0$, in particular $f_{\lambda_i}(0)\to 0$ at the fixed point $x=0$. Hence
        \[
            \lambda_i f(0)\to 0.
        \]
        This is possible for all sequences $\lambda_i\to\infty$ only if
        \[
            f(0)=0.
        \]

        Now we show that $f'(0)=0$. Let $h_i\to 0$ be any sequence with $h_i\neq 0$. For all sufficiently large $i$, we have $h_i\in(-1,1)$. Define
        \[
            \lambda_i=\frac{1}{|h_i|}.
        \]
        Then $\lambda_i\to\infty$, and
        \[
            h_i=\frac{\sgn(h_i)}{\lambda_i}.
        \]
        By the assumption, $f_{\lambda_i}\to 0$ locally uniformly on $\R$. In particular, this convergence is uniform on the compact set $\{-1,1\}$. Therefore,
        \[
            f_{\lambda_i}(\sgn(h_i))\to 0.
        \]
        But
        \[
            f_{\lambda_i}(\sgn(h_i))
            =
            \lambda_i
            f\left(
                \frac{\sgn(h_i)}{\lambda_i}
            \right)
            =
            \lambda_i f(h_i).
        \]
        Hence
        \[
            \lambda_i f(h_i)\to 0.
        \]
        Since $f(0)=0$, we get
        \[
            \left|
                \frac{f(h_i)-f(0)}{h_i}
            \right|
            =
            \left|
                \frac{f(h_i)}{h_i}
            \right|
            =
            \lambda_i |f(h_i)|
            \to 0.
        \]
        Thus
        \[
            \frac{f(h_i)-f(0)}{h_i}\to 0
        \]
        for every sequence $h_i\to 0$ with $h_i\neq 0$. Therefore, $f$ is differentiable at $0$ and
        \[
            f'(0)=0.
        \]

        Therefore, the answer is $\boxed{\text{True}}$.
    \end{solution}

    \begin{problem}{2024.1}
        Determine the number of occurrences of the digit $0$ when all natural numbers from $100$ to $999$ are written consecutively.
    \end{problem}

    \begin{solution}
        Every natural number from $100$ to $999$ is a three-digit number.

        In a three-digit number, the hundreds digit cannot be $0$. Therefore, the digit $0$ can appear only in the tens digit or the units digit.

        First, count the numbers whose tens digit is $0$. The hundreds digit can be one of $1,2,\dots,9$, and the units digit can be one of $0,1,\dots,9$. Hence there are
        \[
            9\cdot 10=90
        \]
        such numbers.

        Next, count the numbers whose units digit is $0$. The hundreds digit can be one of $1,2,\dots,9$, and the tens digit can be one of $0,1,\dots,9$. Hence there are again
        \[
            9\cdot 10=90
        \]
        such numbers.

        Therefore, the total number of appearances of the digit $0$ is
        \[
            90+90=180.
        \]
        Therefore, the answer is $\boxed{180}$.
    \end{solution}

    \begin{problem}{2024.2}
        Determine which of the following people was born latest.
        \[
            \begin{array}{ll}
                (1) & \text{John von Neumann} \\
                (2) & \text{Alan Turing} \\
                (3) & \text{John Nash} \\
                (4) & \text{Pierre de Fermat}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        We compare their years of birth:
        \[
            \begin{array}{ll}
                (1) & \text{John von Neumann: born in } 1903, \\
                (2) & \text{Alan Turing: born in } 1912, \\
                (3) & \text{John Nash: born in } 1928, \\
                (4) & \text{Pierre de Fermat: born in } 1607.
            \end{array}
        \]
        Among these four, the person born the latest is John Nash.

        Therefore, the answer is $\boxed{(3)\ \text{John Nash}}$.
    \end{solution}

    \begin{problem}{2024.3}
        The verified original statement is unavailable. This entry is retained and will be completed when the source is recovered.
    \end{problem}

    \begin{solution}
        The original source is unavailable, so no reliable reconstruction or solution is possible. This entry is retained to preserve the original numbering.
    \end{solution}

    \begin{problem}{2024.4}
        Let $C$ be a convex subset of $\R^3$. Here convex means that for any two points $\mathbf{p},\mathbf{q}\in C$, the line segment joining $\mathbf{p}$ and $\mathbf{q}$ is also contained in $C$.

        Suppose that the following two facts are known:
        \begin{enumerate}
            \item[(i)] The intersection of $C$ with the $xy$-plane is
            \[
                C\cap\{z=0\}
                =
                \{(x,y,0):x^2+4y^2\leq 1\}.
            \]
            \item[(ii)] The set $C$ contains the $z$-axis.
        \end{enumerate}
        Determine the volume of $C$ in the region $-1\leq z\leq 1$, or determine that the given information is insufficient.
        \[
            \begin{array}{ll}
                (1) & \text{This cannot be determined.} \\
                (2) & \pi \\
                (3) & 2\pi \\
                (4) & 4\pi
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Let
        \[
            E=\{(x,y)\in\R^2:x^2+4y^2\leq 1\}.
        \]
        By condition (i),
        \[
            C\cap\{z=0\}=E\times\{0\}.
        \]
        The ellipse $E$ has semiaxes $1$ and $\dfrac{1}{2}$, so its area is
        \[
            \area(E)
            =
            \pi\cdot 1\cdot \frac{1}{2}
            =
            \frac{\pi}{2}.
        \]

        For each height $z$, define the horizontal cross-section
        \[
            C_z
            =
            \{(x,y)\in\R^2:(x,y,z)\in C\}.
        \]
        We claim that $C_z$ has the same area as $E$ for every $z$.

        First, we show that $C_z\subseteq E$. Suppose that $(x,y,z)\in C$ and
        \[
            x^2+4y^2>1.
        \]
        Since $C$ contains the $z$-axis, the point $(0,0,t)$ belongs to $C$ for every $t\in\R$. By convexity, the line segment joining $(x,y,z)$ and $(0,0,t)$ is contained in $C$.

        For any $\alpha\in(0,1)$, we may choose
        \[
            t=-\frac{\alpha z}{1-\alpha}
        \]
        so that this segment intersects the $xy$-plane at the point
        \[
            (\alpha x,\alpha y,0).
        \]
        Since $\alpha$ can be chosen sufficiently close to $1$, we may arrange that
        \[
            (\alpha x)^2+4(\alpha y)^2>1.
        \]
        This would give a point of $C\cap\{z=0\}$ lying outside $E\times\{0\}$, contradicting condition (i). Therefore,
        \[
            C_z\subseteq E
        \]
        for every $z$.

        Conversely, we show that the interior of $E$ is contained in every $C_z$. Take $(x,y)$ with
        \[
            x^2+4y^2<1.
        \]
        Choose $\alpha\in(0,1)$ sufficiently close to $1$ so that
        \[
            \left(\frac{x}{\alpha}\right)^2
            +
            4\left(\frac{y}{\alpha}\right)^2
            \leq 1.
        \]
        Then
        \[
            \left(\frac{x}{\alpha},\frac{y}{\alpha},0\right)\in C.
        \]
        Also, since $C$ contains the $z$-axis,
        \[
            \left(0,0,\frac{z}{1-\alpha}\right)\in C.
        \]
        By convexity, their convex combination also belongs to $C$, and
        \[
            \alpha
            \left(\frac{x}{\alpha},\frac{y}{\alpha},0\right)
            +
            (1-\alpha)
            \left(0,0,\frac{z}{1-\alpha}\right)
            =
            (x,y,z).
        \]
        Hence $(x,y)\in C_z$.

        Thus, for every $z$,
        \[
            \operatorname{int}(E)\subseteq C_z\subseteq E.
        \]
        Since the boundary of the ellipse has area $0$, we have
        \[
            \area(C_z)=\area(E)=\frac{\pi}{2}
        \]
        for every $z$. Therefore, the volume of $C$ in the region $-1\leq z\leq 1$ is
        \[
            \int_{-1}^{1}\area(C_z)\dd z
            =
            \int_{-1}^{1}\frac{\pi}{2}\dd z
            =
            \pi.
        \]

        Therefore, the answer is $\boxed{(2)\ \pi}$.
    \end{solution}

    \begin{problem}{2025.1}
        A fair coin is tossed repeatedly until either two heads in a row or two tails in a row appear. Determine the expected number of tosses before the process stops.
    \end{problem}

    \begin{solution}
        On the first toss, it is impossible to have two consecutive equal outcomes. After the first toss, the process stops exactly when the next toss gives the same result as the previous toss.

        Since the coin is fair, from the second toss onward we have
        \[
            \Prob(\text{the new toss matches the previous toss})
            =
            \frac{1}{2}.
        \]
        Therefore, after the first toss, the number of additional tosses needed until the process stops follows a geometric distribution with success probability $\dfrac{1}{2}$. Its expected value is
        \[
            \frac{1}{1/2}=2.
        \]
        Since the first toss is always made, the expected total number of tosses is
        \[
            1+2=3.
        \]
        Therefore, the answer is $\boxed{3}$.
    \end{solution}

    \begin{problem}{2025.2}
        The nineteenth-century mathematicians Cauchy and Weierstrass established rigorous definitions of limits. Identify the first mathematician to prove that the sum of the reciprocals of the squares of all positive integers is
        \[
            \sum_{n=1}^{\infty}\frac{1}{n^2}
            =
            \frac{\pi^2}{6}.
        \]
        \[
            \begin{array}{ll}
                (1) & \text{David Hilbert} \\
                (2) & \text{Georg Cantor} \\
                (3) & \text{Bernhard Riemann} \\
                (4) & \text{Leonhard Euler}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The problem of finding the exact value of the infinite series
        \[
            \sum_{n=1}^{\infty}\frac{1}{n^2}
        \]
        is known as the Basel problem.

        This problem had been known since the seventeenth century, and it was Leonhard Euler who first found and proved the famous identity
        \[
            \sum_{n=1}^{\infty}\frac{1}{n^2}
            =
            \frac{\pi^2}{6}.
        \]
        Therefore, the answer is $\boxed{(4)\ \text{Leonhard Euler}}$.
    \end{solution}

    \begin{problem}{2025.3}
        Find the equation of the smallest sphere tangent to all four planes
        \[
            x=0,\qquad y=0,\qquad z=0,\qquad x+y+z=3.
        \]
    \end{problem}

    \begin{solution}
        Since the sphere is tangent to the three coordinate planes
        \[
            x=0,\qquad y=0,\qquad z=0,
        \]
        and we are looking for the smallest such sphere, its center must lie in the first octant and have the form
        \[
            \mathbf{c}=(a,a,a)
        \]
        by symmetry. The distance from $\mathbf{c}$ to each coordinate plane is $a$, so the radius of the sphere is
        \[
            r=a.
        \]

        The sphere must also be tangent to the plane
        \[
            x+y+z=3.
        \]
        The distance from $(a,a,a)$ to this plane is
        \[
            \frac{|a+a+a-3|}{\sqrt{1^2+1^2+1^2}}
            =
            \frac{3-3a}{\sqrt{3}},
        \]
        since the smallest sphere lies between the coordinate planes and the plane $x+y+z=3$. This distance must equal the radius $a$, so
        \[
            \frac{3-3a}{\sqrt{3}}=a.
        \]
        Hence
        \[
            3-3a=\sqrt{3}a,
        \]
        and therefore
        \[
            a
            =
            \frac{3}{3+\sqrt{3}}
            =
            \frac{3-\sqrt{3}}{2}.
        \]

        Thus the center is
        \[
            \left(
                \frac{3-\sqrt{3}}{2},
                \frac{3-\sqrt{3}}{2},
                \frac{3-\sqrt{3}}{2}
            \right),
        \]
        and the radius is
        \[
            \frac{3-\sqrt{3}}{2}.
        \]
        Therefore, the equation of the sphere is
        \[
            \boxed{
            \left(x-\frac{3-\sqrt{3}}{2}\right)^2
            +
            \left(y-\frac{3-\sqrt{3}}{2}\right)^2
            +
            \left(z-\frac{3-\sqrt{3}}{2}\right)^2
            =
            \left(\frac{3-\sqrt{3}}{2}\right)^2
            }.
        \]
    \end{solution}

    \begin{problem}{2025.4}
        There are $200$ cards numbered from $1$ to $200$, all initially facing up. For each integer from $1$ to $200$, we call out that integer and flip every card whose number is a multiple of it. Determine the number of cards facing up after all $200$ calls.
        \[
            \begin{array}{ll}
                (1) & 142 \\
                (2) & 153 \\
                (3) & 164 \\
                (4) & 175 \\
                (5) & 186
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        A card numbered $n$ is flipped once for each positive divisor of $n$. Hence it is flipped $\tau(n)$ times, where $\tau(n)$ is the number of positive divisors of $n$.

        A positive integer has an odd number of positive divisors if and only if it is a perfect square. Therefore, exactly the cards numbered
        \[
            1^2,\quad 2^2,\quad 3^2,\quad \ldots,\quad 14^2
        \]
        are flipped an odd number of times, since
        \[
            14^2=196\leq 200<225=15^2.
        \]
        These $14$ cards end up facing down. The remaining
        \[
            200-14=186
        \]
        cards are facing up.

        Therefore, the answer is $\boxed{(5)\ 186}$.
    \end{solution}


    \begin{problem}{2025.5}
        Determine which of $11^9$ and $9^{11}$ is larger.
    \end{problem}

    \begin{solution}
        Taking logarithms, it is enough to compare
        \[
            9\ln 11
            \qquad\text{and}\qquad
            11\ln 9.
        \]
        Equivalently, compare
        \[
            \frac{\ln 11}{11}
            \qquad\text{and}\qquad
            \frac{\ln 9}{9}.
        \]
        The function $f(x)=\ln x/x$ is decreasing for $x>e$. Since $9<11$ and both are greater than $e$, we have
        \[
            \frac{\ln 9}{9}
            >
            \frac{\ln 11}{11}.
        \]
        Hence
        \[
            11\ln 9>9\ln 11,
        \]
        so
        \[
            9^{11}>11^9.
        \]
        Therefore, the larger number is $\boxed{9^{11}}$.
    \end{solution}

    \begin{problem}{2025.6}
        Roll a fair $6$-sided die repeatedly until a $6$ appears. Determine the expected sum of all rolls, including the final $6$.
    \end{problem}

    \begin{solution}
        Let $E$ be the desired expected sum. On the first roll, if the result is $6$, the process stops and contributes $6$. If the result is one of $1,2,3,4,5$, then the process returns to the same situation after contributing that roll.

        Therefore,
        \[
            E
            =
            \frac16\cdot 6
            +
            \frac16\sum_{k=1}^{5}(k+E).
        \]
        Hence
        \[
            E
            =
            1+\frac{15+5E}{6}
            =
            \frac72+\frac56E.
        \]
        Thus
        \[
            \frac16E=\frac72,
            \qquad
            E=21.
        \]
        Therefore, the answer is $\boxed{21}$.
    \end{solution}

    \begin{problem}{2025.7}
        Determine the probability that at least two people in a group of $10$ share the same birthday. Assume a $365$-day year and that each day is equally likely.
    \end{problem}

    \begin{solution}
        It is easier to compute the complementary probability that all $10$ birthdays are distinct. The first person may have any birthday, the second must avoid that birthday, the third must avoid the first two birthdays, and so on. Hence
        \[
            \Prob(\text{all birthdays are distinct})
            =
            \frac{365}{365}\cdot\frac{364}{365}\cdot\frac{363}{365}\cdots\frac{356}{365}.
        \]
        Therefore,
        \[
            \Prob(\text{at least two people share a birthday})
            =
            1-\frac{365}{365}\cdot\frac{364}{365}\cdot\frac{363}{365}\cdots\frac{356}{365}.
        \]
        Therefore, the answer is
        \[
            \boxed{1-\frac{364}{365}\cdot\frac{363}{365}\cdots\frac{356}{365}}.
        \]
    \end{solution}

    \begin{problem}{2025.8}
        Find the largest of the following numbers:
        \[
            \begin{array}{ll}
                (1) & e^\pi \\
                (2) & \pi^e \\
                (3) & \sqrt{500} \\
                (4) & 7\pi
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Using the standard approximations
        \[
            e\approx 2.718,
            \qquad
            \pi\approx 3.142,
        \]
        we get
        \[
            e^\pi \approx 23.14,
            \qquad
            \pi^e \approx 22.46,
            \qquad
            \sqrt{500} \approx 22.36,
            \qquad
            7\pi \approx 21.99.
        \]
        Therefore the largest number is $e^\pi$.

        Therefore, the answer is $\boxed{(1)\ e^\pi}$.
    \end{solution}

    \begin{problem}{2025.9}
        There are three doors. Behind one door is a car, and behind the other two doors are goats. After you choose one door, the host opens one of the remaining doors and shows a goat. You are then given a chance either to keep your original choice or to switch to the other unopened door. Determine which choice gives the higher probability of winning the car.
        \[
            \begin{array}{ll}
                (1) & \text{The probabilities are the same.} \\
                (2) & \text{It is better not to switch.} \\
                (3) & \text{It cannot be determined.} \\
                (4) & \text{It is better to switch.}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The initial choice has probability $1/3$ of being correct and probability $2/3$ of being wrong.

        If the initial choice is correct, switching loses. If the initial choice is wrong, the host must reveal the other wrong door, so switching wins. Therefore switching wins exactly when the initial choice was wrong, which has probability
        \[
            \frac23.
        \]
        Thus switching is better.

        Therefore, the answer is $\boxed{(4)\ \text{It is better to switch}}$.
    \end{solution}

    \begin{problem}{2025.10}
        The verified original statement is unavailable. The surviving record indicates that the problem concerned a real sequence and an infinite series, but the essential formulas are missing. This entry is retained and will be completed when the source is recovered.
    \end{problem}

    \begin{solution}
        The original source is unavailable, so no reliable reconstruction or solution is possible. This entry is retained to preserve the original numbering and will be completed when a verified source becomes available.
    \end{solution}

    \begin{problem}{2025.11}
        Determine which of the following Fields Medalists majored in history rather than mathematics as an undergraduate.
        \[
            \begin{array}{ll}
                (1) & \text{June Huh} \\
                (2) & \text{Caucher Birkar} \\
                (3) & \text{Edward Witten} \\
                (4) & \text{Simon Donaldson} \\
                (5) & \text{William Thurston}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Edward Witten is famous as a mathematical physicist and was the first physicist to receive the Fields Medal. As an undergraduate, he majored in history rather than mathematics.

        Therefore, the answer is $\boxed{(3)\ \text{Edward Witten}}$.
    \end{solution}

    \begin{problem}{2025.12}
        Determine which of the following mathematicians was not alive during the twentieth century.
        \[
            \begin{array}{ll}
                (1) & \text{Sophus Lie} \\
                (2) & \text{Wilhelm Killing} \\
                (3) & \text{Alfred Young} \\
                (4) & \text{Hans Fitting} \\
                (5) & \text{Botong Wang}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Sophus Lie lived from $1842$ to $1899$, so he died before the twentieth century began. The other listed mathematicians were alive at some point during the twentieth century.

        Therefore, the answer is $\boxed{(1)\ \text{Sophus Lie}}$.
    \end{solution}

    \begin{problem}{2025.13}
        K\"onigsberg was the capital of East Prussia and a major center of European scholarship for many years. It was also the hometown of Immanuel Kant, David Hilbert, and J\"urgen Moser. The famous Seven Bridges of K\"onigsberg problem, in which Leonhard Euler proved that no solution exists, marked the beginning of both topology and graph theory. Determine the present-day country containing the former city of K\"onigsberg.
        \[
            \begin{array}{ll}
                (1) & \text{Germany} \\
                (2) & \text{Poland} \\
                (3) & \text{Lithuania} \\
                (4) & \text{Russia} \\
                (5) & \text{Korea}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The historical city K\"onigsberg is now Kaliningrad. Kaliningrad is a city in Russia.

        Therefore, the answer is $\boxed{(4)\ \text{Russia}}$.
    \end{solution}

    \begin{problem}{2025.14}
        You flip a fair coin $50$ times, recording each outcome as either heads $(H)$ or tails $(T)$. A run is a consecutive sequence of the same outcome. For example, if the result is $HHTTTH$, there are $3$ runs: $HH$, $TTT$, and $H$. Determine the expected number of runs.
        \[
            \begin{array}{ll}
                (1) & 24.5 \\
                (2) & 25 \\
                (3) & 25.5 \\
                (4) & 26
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The first toss always starts one run. For each position from the second toss onward, a new run begins exactly when the current toss differs from the previous toss.

        For a fair coin,
        \[
            \Prob(\text{current toss differs from previous toss})=\frac12.
        \]
        Let $I_i$ be the indicator that a new run starts at position $i$, for $2\leq i\leq 50$. Then
        \[
            R=1+\sum_{i=2}^{50}I_i.
        \]
        By linearity of expectation,
        \[
            \Exp[R]
            =
            1+\sum_{i=2}^{50}\Exp[I_i]
            =
            1+49\cdot\frac12
            =
            25.5.
        \]
        Therefore, the answer is $\boxed{(3)\ 25.5}$.
    \end{solution}

    \begin{problem}{2025.15}
        Determine which of the following theorems was published first.
        \[
            \begin{array}{ll}
                (1) & \text{Bayes' Theorem} \\
                (2) & \text{Intermediate Value Theorem} \\
                (3) & \text{Cauchy--Schwarz Inequality} \\
                (4) & \text{Fundamental Theorem of Calculus}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The Fundamental Theorem of Calculus is associated with Newton and Leibniz in the seventeenth century, while Bayes' Theorem, the Intermediate Value Theorem in its modern rigorous form, and the Cauchy--Schwarz Inequality were published later.

        Therefore, the earliest among the choices is the Fundamental Theorem of Calculus.

        Therefore, the answer is $\boxed{(4)\ \text{Fundamental Theorem of Calculus}}$.
    \end{solution}

    \begin{problem}{2025.16}
        Determine which of the following amounts is closest to the Abel Prize award.
        \[
            \begin{array}{ll}
                (1) & \text{None} \\
                (2) & \text{About }\$1{,}000{,}000 \\
                (3) & \text{About }\$1{,}500{,}000 \\
                (4) & \text{About }\$2{,}000{,}000
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        The Abel Prize is a major international mathematics prize awarded by the Norwegian Academy of Science and Letters. Its prize money is commonly described, at quiz-level precision, as being closest to about one million US dollars among the listed options.

        Therefore, the answer is $\boxed{(2)\ \text{About }\$1{,}000{,}000}$.
    \end{solution}

    \begin{problem}{2025.17}
        Determine which of the following mathematicians has received both the Fields Medal and the Abel Prize.
        \[
            \begin{array}{ll}
                (1) & \text{John Milnor} \\
                (2) & \text{Terence Tao} \\
                (3) & \text{June Huh} \\
                (4) & \text{No one}
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        John Milnor received the Fields Medal in $1962$ and the Abel Prize in $2011$. Therefore he has received both prizes.

        Therefore, the answer is $\boxed{(1)\ \text{John Milnor}}$.
    \end{solution}

    \begin{problem}{2025.18}
        Suppose you roll a fair $6$-sided die until you have seen all $6$ faces. Determine the probability that no odd-numbered face appears before all even-numbered faces have appeared.
        \[
            \begin{array}{ll}
                (1) & 20\% \\
                (2) & 15\% \\
                (3) & 10\% \\
                (4) & 5\%
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Ignore repeated rolls and look only at the order in which new faces appear. The condition is that the first three distinct faces to appear are exactly the even faces $2,4,6$ in some order.

        The probability that the first new face is even is $3/6$. Given that one even face has appeared, the probability that the next new face is also even is $2/5$. Given that two even faces have appeared, the probability that the next new face is the remaining even face is $1/4$. Therefore the desired probability is
        \[
            \frac36\cdot\frac25\cdot\frac14
            =
            \frac1{20}
            =
            5\%.
        \]
        Therefore, the answer is $\boxed{(4)\ 5\%}$.
    \end{solution}

    \begin{problem}{2025.19}
        In the Tower of Hanoi puzzle, $5$ disks are stacked on peg $A$ from largest to smallest. We want to move all disks to peg $C$ using auxiliary peg $B$, subject to the following rules:
        \begin{itemize}
            \item Only one disk may be moved at a time.
            \item A larger disk may not be placed on a smaller disk.
            \item Peg $B$ may be used freely.
        \end{itemize}
        Determine the minimum number of moves required.
        \[
            \begin{array}{ll}
                (1) & 25 \\
                (2) & 31 \\
                (3) & 32 \\
                (4) & 63 \\
                (5) & 120
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Let $T(n)$ be the minimum number of moves needed for $n$ disks. To move $n$ disks, first move the top $n-1$ disks to the auxiliary peg, then move the largest disk once, and finally move the $n-1$ disks onto the largest disk. Hence
        \[
            T(n)=2T(n-1)+1,
            \qquad
            T(1)=1.
        \]
        Solving this recurrence gives
        \[
            T(n)=2^n-1.
        \]
        Therefore,
        \[
            T(5)=2^5-1=31.
        \]
        Therefore, the answer is $\boxed{(2)\ 31}$.
    \end{solution}

    \begin{problem}{2025.20}
        Determine the last two digits of $7^{123}$, expressed as a two-digit number.
        \[
            \begin{array}{ll}
                (1) & 37 \\
                (2) & 41 \\
                (3) & 43 \\
                (4) & 49
            \end{array}
        \]
    \end{problem}

    \begin{solution}
        Work modulo $100$. Since
        \[
            7^4=2401\equiv 1\pmod{100},
        \]
        and
        \[
            123=4\cdot 30+3,
        \]
        we have
        \[
            7^{123}
            =
            (7^4)^{30}7^3
            \equiv
            1^{30}\cdot 343
            \equiv
            43
            \pmod{100}.
        \]
        Therefore, the last two digits are $43$.

        Therefore, the answer is $\boxed{(3)\ 43}$.
    \end{solution}
