    \chapter*{Preface}
    \markboth{Preface}{Preface}
    \addcontentsline{toc}{chapter}{Preface}

    This book is devoted to the Mathematics section of Science Quiz in the
    \POSTECH--\KAIST{} Science War.
    It collects mathematical concepts, facts to memorize, past problems, and
    expected problems that are useful for competition preparation.

    Science Quiz remains the primary target of this book. At the same time,
    selected results and techniques from mathematical olympiads, gifted-school
    and science-high-school entrance examinations, and other problem-solving
    competitions are included when they are concise, memorable, or broadly
    useful. The emphasis is therefore on meaningful results, effective formulas,
    and decisive ideas rather than on reproducing long proofs.

    The main purpose of this text is not merely to record solutions, but to
    build a practical and reliable mathematical toolkit for Science Quiz.
    It includes standard theorems, important terminology, memorable facts about
    mathematicians and mathematical history, and problem-solving strategies that
    may be useful in an actual match.

    Science Quiz requires competitors to move rapidly across broad areas of
    mathematics and solve each problem in roughly one or two minutes. Accordingly,
    this book combines the academic foundations of university mathematics with
    contest-style tools such as invariants, extremal arguments, coloring, symmetry,
    reverse engineering, fast transformations, and pattern recognition. The aim is
    not to reproduce long olympiad proofs, but to make the decisive idea or
    calculation available quickly when a short and creative problem appears.

    I sincerely hope that this book contributes to \POSTECH's victory.
    Good luck, and let the game begin.
